# Cohort Summary Table

## Total number of samples

Firstly, look at the total number of samples:

```{r}
#recur variables here
avtum=table(Cdata$Rat_ID)
Cchar=Cdata[Cdata$Cohort!="Progression", ]
Cprog=Cdata[Cdata$Cohort=="Progression", ]
```

Feature | Levels | N
--------|---------|-----
Total No Tumors |All | `r nrow(Cdata)`
- | Characterisation| `r length(which(Cdata$Cohort=="Characterisation"))`
- | Progression | `r length(which(Cdata$Cohort=="Progression"))`
Total No Rats | All| `r length(unique(Cdata$Rat_ID))`
- | Characterisation| `r length(unique(Cdata$Rat_ID[which(Cdata$Cohort=="Characterisation")]))`
- | Progression | `r length(unique(Cdata$Rat_ID[which(Cdata$Cohort=="Progression")]))`
No tumours per rat | | `r round(mean(avtum),2) ` ( `r quantile(avtum, c(0.1, 0.9))`)

## Compare the characterisation vs progression cohort

Feature | Levels | Characterisation | Progression
--------|---------|-------- |------
Total No tumours | | `r nrow(Cchar)` | `r nrow(Cprog)`
Treatments| Vehicle | `r length(which(Cdata$Treatment=="Vehicle"))`| `r length(which(Cprog$Treatment=="Vehicle"))`
-| LY | | `r length(which(Cprog$Treatment=="LY"))`
-| PDL1| | `r length(which(Cprog$Treatment=="PDL1"))`
-| PDL1+LY| | `r length(which(Cprog$Treatment=="PDL1+LY"))`
- | NA | `r length(which(is.na(Cchar$Treatment)))` |`r length(which(is.na(Cprog$Treatment)))`
Histology |  diff. adenocarcinomas | `r length(which(Cchar$Tumor.Histology=="well differentiated mammary solid adenocarcinoma"))`|`r length(which(Cprog$Tumor.Histology=="well differentiated mammary solid adenocarcinoma"))`
- | mucinous carcinoma | `r length(which(Cchar$Tumor.Histology=="mucinous carcinoma"))`|`r length(which(Cprog$Tumor.Histology=="mucinous carcinoma"))`
- | Fibroadenoma | `r length(which(Cchar$Tumor.Histology=="fibroadenoma"))`|`r length(which(Cprog$Tumor.Histology=="fibroadenoma"))`
- | NA |`r length(which(is.na(Cchar$Tumor.Histology)))` | `r length(which(is.na(Cprog$Tumor.Histology)))`
Age Injection | 32-36 | `r length(which(Cchar$Age.at.injection=="32-36")) `|`r length(which(Cprog$Age.at.injection=="32-36")) `
 - | 35 | `r length(which(Cchar$Age.at.injection=="35")) ` | `r length(which(Cprog$Age.at.injection=="35")) `
  - | 49 | `r length(which(Cchar$Age.at.injection=="49")) `|`r length(which(Cprog$Age.at.injection=="49")) `
 - | NA | `r length(which(is.na(Cchar$Age.at.injection))) `|`r length(which(is.na(Cprog$Age.at.injection))) `
Time (days) | NMU 2 Sac| `r round(mean(Cchar$Time.NMU2Sac, na.rm=T),2) ` (`r quantile(Cchar$Time.NMU2Sac, c(0.1, 0.9), na.rm=T)`) | `r round(mean(Cprog$Time.NMU2Sac, na.rm=T), 2) ` (`r quantile(Cprog$Time.NMU2Sac, c(0.1, 0.9), na.rm=T)`)
 - | Cases with NA| `r length(which(is.na(Cchar$Time.NMU2Sac)))` | `r length(which(is.na(Cprog$Time.NMU2Sac)))` 
 - | NMU 2 Tumor | | `r round(mean(Cprog$Time.NMU2Tumor, na.rm=T), 2)` (`r  quantile(Cprog$Time.NMU2Tumor, c(0.1, 0.9), na.rm=T)`)
 - | Cases with NA| `r length(which(is.na(Cchar$Time.NMU2Tumor)))` | `r length(which(is.na(Cprog$Time.NMU2Tumor)))` 
 - | Tum Spec Surv|  | `r round(mean(Cprog$Time.Tumor2Sac, na.rm=T),2)` (`r quantile(Cprog$Time.Tumor2Sac, c(0.1, 0.9), na.rm=T)`)
 - | Cases with NA| `r length(which(is.na(Cchar$Time.Tumor2Sac)))` | `r length(which(is.na(Cprog$Time.Tumor2Sac)))`
Growth Rate/Size (mm) | overall size @ sac | `r round(mean(Cchar$Tumor.diameter.sac.mm,na.rm=T), 2)`  (`r quantile(Cchar$Tumor.diameter.sac.mm, c(0.1,  0.9), na.rm=T)`) | `r round(mean(Cprog$Tumor.diameter.sac.mm,na.rm=T), 2)` (`r quantile(Cprog$Tumor.diameter.sac.mm, c(0.1, 0.9), na.rm=T)`)
- | Growing No. | | `r length(grep("growing", Cprog$Tumor.Growth))`
- | Growing size @ sac |  |  `r round(mean(Cprog$Tumor.diameter.sac.mm[grep("growing", Cprog$Tumor.Growth)]),2)` `r quantile(Cprog$Tumor.diameter.sac.mm[grep("growing", Cprog$Tumor.Growth)], c(0.1, 0.5, 0.9), na.rm=T)`
- | Stable ||`r length(grep("stable", Cprog$Tumor.Growth))`
- | Stable size @ sac | | `r round(mean(Cprog$Tumor.diameter.sac.mm[grep("stable", Cprog$Tumor.Growth)]),2)` (`r quantile(Cprog$Tumor.diameter.sac.mm[grep("stable", Cprog$Tumor.Growth)], c(0.1, 0.5, 0.9), na.rm=T)`)
- | NA |`r length(which(is.na(Cchar$Tumor.Growth)))` | `r length(which(is.na(Cprog$Tumor.Growth)))`
Spatial Pattern| | |
- | Infiltrating| `r length(which(Cchar$Pathologist.CD8.Spatial=="Infiltrating"))` |  `r length(which(Cprog$Pathologist.CD8.Spatial=="Infiltrating"))`
- | Restricted |  `r length(which(Cchar$Pathologist.CD8.Spatial=="restricted"))` |  `r length(which(Cprog$Pathologist.CD8.Spatial=="restricted"))`
- | NA| `r length(which(is.na(Cchar$Pathologist.CD8.Spatial)))` |  `r length(which(is.na(Cprog$Pathologist.CD8.Spatial)))`
RNA samples | any fraction | `r length(which(!is.na(Cchar$AnyRdata)))`| `r length(which(!is.na(Cprog$AnyRdata)))`
- | Ep|`r length(which(!is.na(Cchar$FqFile.Ep)))`|`r length(which(!is.na(Cprog$FqFile.Ep)))`
- | DN|`r length(which(!is.na(Cchar$FqFile.DN)))`|`r length(which(!is.na(Cprog$FqFile.DN)))`
- | CD45| `r length(which(!is.na(Cchar$FqFile.CD45)))`| `r length(which(!is.na(Cprog$FqFile.CD45)))`
Imaging Data | No tumors | | `r length(which(!is.na(Cprog$WSI.Image.me)))`
- | No tumors with RNA | |`r length(which(!is.na(Cprog$WSI.Image.me)& !is.na(Cprog$AnyRdataD)))`
Trichrome Data | No tumors | |`r length(which(!is.na(Cprog$Trichrome)))`
- | No tumors with RNA | |`r length(which(!is.na(Cprog$Trichrome)& !is.na(Cprog$AnyRdataD)))`
FACS data | Comprehensive|  | `r length(which(colnames(Fdata)%in%gsub("_", "", Cprog$TumorID)))`
- | EpCAM/CD45 | `r length(which(!is.na(Cchar$CD45.Frac.FACS)))` |  `r length(which(!is.na(Cprog$CD45.Frac.FACS)))` 

## Summary of the RNA data

Below is a table of the samples with RNA information

```{r}
CRchar=Cchar[which(Cchar$AnyRdata=="yes"), ]
CRprog=Cprog[which(Cprog$AnyRdata=="yes"), ]
```

Feature | Levels | Characterisation | Progression
--------|---------|-------- |------
RNA samples | any fraction | `r length(which(!is.na(CRchar$RNA.seqSampleID)))`| `r length(which(!is.na(CRprog$RNA.seqSampleID)))`
- | Ep|`r length(which(!is.na(CRchar$FqFile.Ep)))`|`r length(which(!is.na(CRprog$FqFile.Ep)))`
- | DN|`r length(which(!is.na(CRchar$FqFile.DN)))`|`r length(which(!is.na(CRprog$FqFile.DN)))`
- | CD45| `r length(which(!is.na(CRchar$FqFile.CD45)))`| `r length(which(!is.na(CRprog$FqFile.CD45)))`
Treatments Char/Prev  | Vehicle | `r length(which(CRchar$Treatment=="Vehicle"))`| `r length(which(CRprog$Treatment=="Vehicle"))`
- | Untreat (char) | `r length(which(CRchar$Treatment=="Untreated"))` |
-| LY | | `r length(which(CRprog$Treatment=="LY"))`
-| PDL1| | `r length(which(CRprog$Treatment=="PDL1"))`
-| PDL1+LY| | `r length(which(CRprog$Treatment=="PDL1+LY"))`
- | NA | `r length(which(is.na(CRchar$Treatment) ))` |`r length(which(is.na(CRprog$Treatment)))`
Time | NMU 2 Sac| `r round(mean(CRchar$Time.NMU2Sac, na.rm=T),2) ` (`r quantile(CRchar$Time.NMU2Sac, c(0.1, 0.9), na.rm=T)`) | `r round(mean(CRprog$Time.NMU2Sac, na.rm=T), 2) ` (`r quantile(CRprog$Time.NMU2Sac, c(0.1, 0.9), na.rm=T)`)
Growth Rate/Size (mm) | overall size @ sac | `r round(mean(CRchar$Tumor.diameter.sac.mm,na.rm=T), 2)`  (`r quantile(CRchar$Tumor.diameter.sac.mm, c(0.1,  0.9), na.rm=T)`) | `r round(mean(CRprog$Tumor.diameter.sac.mm,na.rm=T), 2)` (`r quantile(CRprog$Tumor.diameter.sac.mm, c(0.1, 0.9), na.rm=T)`)
- | Growing No. | | `r length(grep("growing", CRprog$Tumor.Growth))`
- | Growing size @ sac |  |  `r round(mean(CRprog$Tumor.diameter.sac.mm[grep("growing", CRprog$Tumor.Growth)]),2)` `r quantile(CRprog$Tumor.diameter.sac.mm[grep("growing", CRprog$Tumor.Growth)], c(0.1, 0.5, 0.9), na.rm=T)`
- | Stable ||`r length(grep("stable", CRprog$Tumor.Growth))`
- | Stable size @ sac | | `r round(mean(CRprog$Tumor.diameter.sac.mm[grep("stable", CRprog$Tumor.Growth)]),2)` (`r quantile(CRprog$Tumor.diameter.sac.mm[grep("stable", CRprog$Tumor.Growth)], c(0.1, 0.5, 0.9), na.rm=T)`)
- | NA |`r length(which(is.na(CRchar$Tumor.Growth)))` | `r length(which(is.na(CRprog$Tumor.Growth)))`
Growth and Treatment: comparing small/stable vs large/growing | | 
Vehicle | N s/l |  | `r length(which(CRprog$Treatment=="Vehicle" & CRprog$Tumor.Growth=="stable"))`, `r length(which(CRprog$Treatment=="Vehicle" & CRprog$Tumor.Growth=="growing"))`
LY | N s/l |  | `r length(which(CRprog$Treatment=="LY" & CRprog$Tumor.Growth=="stable"))`, `r length(which(CRprog$Treatment=="LY" & CRprog$Tumor.Growth=="growing"))`
PDL1 and Treatment | N s/l |  | `r length(which(CRprog$Treatment=="PDL1" & CRprog$Tumor.Growth=="stable"))`, `r length(which(CRprog$Treatment=="PDL1" & CRprog$Tumor.Growth=="growing"))`
PDL1+LY | N s/l |  | `r length(which(CRprog$Treatment=="PDL1+LY" & CRprog$Tumor.Growth=="stable"))`, `r length(which(CRprog$Treatment=="PDL1+LY" & CRprog$Tumor.Growth=="growing"))`


