# DESeq analysis: Immunotherapy/Growth comparisons

## Summary of comparisons

Quick plot of the differences in number of differential genes:

```{r}
Allchanges=rbind(UpDn1, CDUpDn1, DNUpDn1)
Allchanges=cbind(Allchanges, Frac=c("Ep", "Ep", "CD", "CD", "DN", "DN"), dir=rownames(Allchanges))

AllchangesB=melt(as.data.frame(Allchanges), id.vars = c("Frac", "dir"))

ggplot(AllchangesB, aes(x=variable, y=as.numeric(value), fill=dir))+geom_bar(stat="identity",position = "dodge")+facet_grid(Frac~., scale="free_y", space="free_y")+theme(axis.text.x = element_text(angle = 90, hjust = 1))+ylab("No. differential genes")+ggtitle("number of differential genes")+scale_y_log10()
## plot the number of samples in each comparison
```

Notes:

Comparison Type | Main cell type impacted
----------------|-----------------------
Treatment | CD45+, PDL1+LY in the DN, LY may affect Ep
growth | Ep and CD45
growth in control | CD and DN
Spatial Patterns | CD45, Ep
Spatial, high cd8 | all cases


## DESeq: any treatment vs control

What are the differential genes if we compare any immunotherapy to the control arm?

First look at the epithelial fraction:

```{r}
# EpddsTreatB=EpddsTreat
# design(EpddsTreatB)=~treatA+factor(Batch)
# EpddsTreatB=DESeq(EpddsTreatB)

epr=results(Epdds)
Xa1=epr[which(epr$padj<0.05 & epr$baseMean>60), ]

epvsd=vst(Epdds)
assay(epvsd)<- limma::removeBatchEffect(assay(epvsd),factor(epvsd$Batch))


Mat2=epvsd[match(rownames(Xa1), rownames(epvsd)), ]

colSCols=ColMerge[ ,1][Mat2$Treatment]

heatmap.2(assay(Mat2), ColSideColors = colSCols, trace="none", scale="row", col=RdBu[11:1], hclustfun = hclust.ave)

write.csv(epr, file="nature-tables/Supp2_ep_io_vs_nonio.csv")

```

DN

List of markers from xcell:

```{r}
xcellgenes=readxl::read_xlsx("../anntotations/xcell_genes.xlsx", sheet=1)
AllGenes=as.vector(xcellgenes[ ,-c(1:2)])
AllGenes=firstup(tolower(unique(as.character(unlist(AllGenes)))))

#rm list
rmList=c("15L_B_DN", "6R_B_DN", "10L_C_DN", "6R_D_DN")

DNr=results(DNdds)
Xa1=DNr[which(DNr$padj<0.05 & DNr$baseMean>100 & abs(DNr$log2FoldChange)>1.5), ]

# DNvsd=vst(DNddsTreatB)
# assay(DNvsd)<- limma::removeBatchEffect(assay(DNvsd),factor(DNvsd$Batch))
AllG2=rownames(Xa1)[which(rownames(Xa1)%in%c(AllGenes, "Il6"))]

Mat2=vsdLimmaDN[match(AllG2, rownames(vsdLimmaDN)), ]
colnames(Mat2)=Cdata$NewID[match(substr(colnames(Mat2),1 , nchar(colnames(Mat2))-3), Cdata$TumorID)]

colSCols=ColMerge[ ,1][Mat2$Treatment]

heatmap.2(assay(Mat2), ColSideColors = colSCols, trace="none", scale="row", col=RdBu[11:1])

write.csv(DNr, file="nature-tables/Supp2_DN_io_vs_non_io.csv")

ax1=results(DNddsTreat, c("Treatment", "LY", "Vehicle"))
bx1=results(DNddsTreat, c("Treatment", "PDL1", "Vehicle"))
cx1=results(DNddsTreat, c("Treatment", "PDL1+LY", "Vehicle"))

testout=cbind(ax1, bx1, cx1)
write.csv(testout, file="nature-tables/Supp3_dn_all_pairwise_treatments.csv")


```

CD45

```{r 3j, fig.cap="CD45 IO vs non-IO"}
# CDddsTreatB=CDddsTreat
# design(CDddsTreatB)=~treatA
# CDddsTreatB=DESeq(CDddsTreatB)

CDr=results(CDdds)
Xa1=CDr[which(CDr$padj<0.05 & CDr$baseMean>100 & abs(CDr$log2FoldChange)>1.5), ]

AllG2=rownames(Xa1)[which(rownames(Xa1)%in%RatAllImm)]

Mat2=vsdLimmaCD[match(AllG2, rownames(vsdLimmaCD)), ]
colnames(Mat2)=Cdata$NewID[match(substr(colnames(Mat2),1 , nchar(colnames(Mat2))-5), Cdata$TumorID)]

colSCols=ColMerge[ ,1][Mat2$Treatment]

#pdf("figure-outputs/FIgure3j_1_vs_all.pdf", height=7, width=6)

heatmap.2(assay(Mat2), ColSideColors = colSCols, trace="none", scale="row", col=RdBu[11:1],
          hclustfun = hclust.ave)

#dev.off()

write.csv(assay(Mat2), file="nature-tables/Figure3j_1_vs_all.csv")
write.csv(CDr, file="nature-tables/Supp3_cd45_io_vs_nonio.csv")
```



```{r 3jb, fig.cap="IO vs any other treatment"}
## ALSO LOOK AT the pairwise comparisons and pull out those genes
ax1=results(CDddsTreat, c("Treatment", "LY", "Vehicle"))
bx1=results(CDddsTreat, c("Treatment", "PDL1", "Vehicle"))
cx1=results(CDddsTreat, c("Treatment", "PDL1+LY", "Vehicle"))
xa=rownames(ax1)
xb=rownames(bx1)
xc=rownames(cx1)

G1=rownames(ax1)[which(ax1$baseMean>100 & ax1$padj<0.05 & abs(ax1$log2FoldChange)>2)]
G2=rownames(bx1)[which(bx1$baseMean>100 & bx1$padj<0.05 & abs(bx1$log2FoldChange)>2)]
G3=rownames(cx1)[which(cx1$baseMean>100 & cx1$padj<0.05 & abs(cx1$log2FoldChange)>2)]

AllG=unique(c(G1, G2, G3))

AllG2=AllG[which(AllG%in%RatAllImm)]

Mat2=vsdLimmaCD[match(AllG2, rownames(vsdLimmaCD)), ]
colnames(Mat2)=Cdata$NewID[match(substr(colnames(Mat2),1 , nchar(colnames(Mat2))-5), Cdata$TumorID)]

colSCols=ColMerge[ ,1][Mat2$Treatment]
#pdf("figure-outputs/FIgure3j_all_comparisons.pdf", height=9, width=6)

heatmap.2(assay(Mat2), ColSideColors = colSCols, trace="none", scale="row", col=RdBu[11:1], hclustfun = hclust.ave)
#dev.off()

testout=cbind(ax1, bx1, cx1)
write.csv(testout, file="nature-tables/Supp3_cd45_all_pairwise_treatments.csv")

xa[100:105]
xb[100:105]
xc[100:105]

#write.csv(assay(Mat2), file="nature-tables/Figure3j_all_comparisons.csv")
```

### EPCAM samples

Note that there are a subset of samples different from the other based on CD8, growth and MH index (6RB, 3NB, 10LD, 8LD, 11ND) but these parameters don't group the other samples well


There is a small subset of samples which have high expression of CCLs and based on the top variable genes, there seems to be three types of epithelial cases. 

```{r, eval=F}
SampleFrac="Ep"

allEpGenes=unlist(EpCompSig)
allEpTab=table(allEpGenes)
EpGenesList=names(allEpTab)[which(allEpTab>1)]
#EpGenesList=EpGenesList[-grep("LOC", EpGenesList)]
#EpGenesList=EpGenesList[-grep("RGD", EpGenesList)]
# also, remove all samples which have a base mean less than 10
x1=EpComp4[[1]]$baseMean[match(EpGenesList, EpComp4[[1]]$Gene)]
x2=which(x1>100)
vstEp=vst(Epdds, blind=T)

#EpddsTreatG$Comp8_v2=factor(paste(EpddsTreatG$Treatment, EpddsTreatG$Growth))

ColsideCols=hue_pal()(2)[Epdds$CD8FracCut]
ColsideColsB=hue_pal()(4)[Epdds$Treatment]
ColsideColsC=hue_pal()(2)[Epdds$Growth]
ColsideColsD=hue_pal()(2)[Epdds$MHcut]

colTable=rbind(ColsideCols, ColsideColsB, ColsideColsC, ColsideColsD)
rownames(colTable)=c("CD8", "Treatment", "growth", "MH")

#pdf(sprintf("rslt/DESeq/DEseq_comparisons_%s.pdf", SampleFrac), width=7, height=10)

sampleDist=dist(t(assay(vstEp)))
sampleDistMatrix <- as.matrix(sampleDist)
heatmap.plus(sampleDistMatrix, trace="none", scale="none", col=brewer.pal(9, "Blues")[9:1], ColSideColors = t(colTable), main=sprintf("Similarity between %s samples", SampleFrac))

# Take the top 150 variable genes 
# compute the sds for each genes
sds = rowSds(assay(vstEp))
means=rowMeans(assay(vstEp))
CVs=sds/means
CVs=CVs[-which(means<7)]
sds2=sort(CVs, decreasing = T)

heatmap.plus(assay(vstEp)[names(sds2)[1:100], ], trace="none", scale="row", col=RdBu[11:1], ColSideColors = t(colTable), main=sprintf("top 100 variable genes all %s samples", SampleFrac))


heatmap.plus(assay(vstEp)[EpGenesList[x2], ], trace="none", scale="row", col=RdBu[11:1], ColSideColors = t(colTable),
          main=sprintf("DEG at least 2 %s samples", SampleFrac))

# Make a table for these genes
FCvalues=matrix(NA, nrow=length(x2), ncol=length(EpComp4))

for (i in 1:length(EpComp4)){
  FCvalues[ ,i]=EpComp4[[i]]$log2FoldChange[match(EpGenesList[x2], EpComp4[[i]]$Gene)]
}

colnames(FCvalues)=names(EpComp4)
rownames(FCvalues)=EpGenesList[x2]


# for each list pick the top 50 diff genes FC>2 and baseman > 50
glist=NA
for (i in 1:length(EpComp4)){
  glist=c(glist, EpComp4[[i]]$Gene[which(EpComp4[[i]]$padj<0.1 & abs(EpComp4[[i]]$log2FoldChange)>2 & EpComp4[[i]]$baseMean>25)])
}

ColsideCols=SetCols[EpddsTreatG$Comp8]

glistmod=unique(glist)
glistmod=glistmod[-grep("LOC", glistmod)]
glistmod=na.omit(glistmod[-grep("RGD", glistmod)])

#heatmap.plus(assay(vstEp)[glistmod, ], trace="none", scale="row", col=RdBu[11:1], ColSideColors = t(colTable),  main=sprintf("All DEG %s samples", SampleFrac))
```

We can also view this in the form of what is highly expressed in each comparison. Below is a heatmap of the fold changes:

```{r, eval=F}
# distance of the samples to one another
par(oma=c(1,0,0,0))
heatmap.2(FCvalues, trace="none", scale="row", col=RdBu[11:1], 
          main=sprintf("All comparisons %s, log2FC for each gene", SampleFrac))
#dev.off()
```

Below, we visualise the enrichment scores using different gene sets and 

```{r, eval=F}
## GSEA results temp.
pdf("~/Desktop/Fig5Gene_set_enrichment.pdf", height=10, width=10)

#Epismr3=ismr4

#pdf(sprintf("rslt/DESeq/GSEA_signature_comparisons_%s.pdf", SampleFrac), width=10, height=12)

for (i in 1:length(Epismr3)){
  x1=Epismr3[[i]]
  # find padj<0.1
  idx2=lapply(seq(3, ncol(x1), by=2), function(x) which(as.numeric(as.character(x1[ ,x]))<0.01))
  unList=table(unlist(idx2))
  #pickThese=names(unList)#[which(unList>1)]
  pickThese2=names(unList)[which(unList>1)]
  
  if (length(pickThese2)>100){
    temp=x1[as.numeric(pickThese2), seq(2, ncol(x1), by=2)]
    temp2=sapply(temp, function(x) as.numeric(as.character(x)))
    temp2=rowSds(abs(temp2))
    names(temp2)=pickThese2
    temp2=sort(temp2, decreasing = T)
    pickThese2=names(temp2)[1:100]
  }
  
  x23=x1[as.numeric(pickThese2), seq(2, ncol(x1), by=2)]
  
  x234=sapply(x23, function(x) as.numeric(as.character(x)))
  rownames(x234)=rownames(x23)
  
  par(oma=c(2,0,0,2))
  heatmap.2(x234, col=RdBu[11:1], scale="none", trace="none", main=sprintf("%s pathways %s fraction", 
                                                                           names(Epismr3)[i], SampleFrac))
}

dev.off()
```

## Growing vs stable emphasis

Below, we focus specifically on the growing vs stable comparison in greater depth.
Here, we look at the 3 different fractions in greater depth and look at volcano plots of DEG and heatmaps of differential genes. Samples are colored according to whether they are growing or stable

### DN fraction

```{r}

#pdf(sprintf("rslt/DESeq/volcano_plots_differences_stable_vs_growing_%s.pdf", Sys.Date()), height=7, width=8)

DNa1=DNComp4$stable_vs_growing
DNa=DNComp4$stable_vs_growing[ which(DNComp4$stable_vs_growing$padj<0.1 & DNComp4$stable_vs_growing$baseMean>50 & abs(DNComp4$stable_vs_growing$log2FoldChange)>1.5), ]
with(DNa1, plot(log2FoldChange, -log10(padj), pch=20, main="Volcano plot: DN Stable (+) vs Growing (-)", cex=1.0, xlab=bquote(~Log[2]~fold~change), ylab=bquote(~-log[10]~Q~value)))
with(subset(DNa1, padj<0.1 & abs(log2FoldChange)>1.5), points(log2FoldChange, -log10(padj), pch=20, col="red", cex=0.5))
with(subset(DNa1, padj<0.1 & abs(log2FoldChange)>1.5), text(log2FoldChange+0.05, -log10(padj)+0.05, Gene, pch=20, col="red", cex=0.75))


heatmap.2(assay(vstDN)[ DNa$Gene, ], scale="row", trace="none", ColSideColors = as.character(as.numeric(vstDN$Growth)), col=RdBu[11:1], main="DN genes")

write.csv(DNa1, file="nature-tables/Supp_2_DEG_DN_growing_stable.csv")
```

### CD fraction

Volcano plot
```{r 4g, fig.cap="volcano plot of DEG stable vs growing CD45"}

#pdf("~/Desktop/Figure4G_CD_stable_growing.pdf", height=12, width=12)
CDa1=CDComp4$stable_vs_growing
CDa=CDComp4$stable_vs_growing[ which(CDComp4$stable_vs_growing$padj<0.05 & CDComp4$stable_vs_growing$baseMean>100 & abs(CDComp4$stable_vs_growing$log2FoldChange)>2), ]
with(CDa1, plot(log2FoldChange, -log10(padj), pch=20, main="Volcano plot: CD Stable (+) vs Growing (-)", cex=1.0, xlab=bquote(~Log[2]~fold~change), ylab=bquote(~-log[10]~Q~value)))
with(subset(CDa1, padj<0.1 & abs(log2FoldChange)>1.5), points(log2FoldChange, -log10(padj), pch=20, col="red", cex=0.5))
with(subset(CDa1, padj<0.1 & abs(log2FoldChange)>1.5), text(log2FoldChange+0.05, -log10(padj)+0.05, Gene, pch=20, col="red", cex=0.75))

write.csv(CDa1, file="nature-tables/4g.csv")
```

Here are a bunch of heatmaps, which are then separated by wehther genes are downregulated (growing specific) or upregulated (stable specific)

```{r }
heatmap.2(assay(vstCD)[ CDa$Gene, ], scale="row", trace="none", ColSideColors = as.character(as.numeric(vstCD$Growth)), col=RdBu[11:1], main="CD genes")

## growing specific
CDg=CDa$Gene[which(CDa$log2FoldChange<0)]
heatmap.2(assay(vstCD)[CDg, ], scale="row", trace="none", ColSideColors = as.character(as.numeric(vstCD$Growth)), col=RdBu[11:1], main="CD genes growing")

## stable specific
CDs=CDa$Gene[which(CDa$log2FoldChange>0)]
heatmap.2(assay(vstCD)[CDs, ], scale="row", trace="none", ColSideColors = as.character(as.numeric(vstCD$Growth)), col=RdBu[11:1], main="CD genes stable")

#dev.off()
```

### Ep Fraction

```{r}
Epa1=EpComp4$stable_vs_growing
Epa=EpComp4$stable_vs_growing[ which(EpComp4$stable_vs_growing$padj<0.1 & EpComp4$stable_vs_growing$baseMean>50 & abs(EpComp4$stable_vs_growing$log2FoldChange)>1.5), ]
with(Epa1, plot(log2FoldChange, -log10(padj), pch=20, main="Volcano plot: Ep Stable vs Growing", cex=1.0, xlab=bquote(~Log[2]~fold~change), ylab=bquote(~-log[10]~Q~value)))
with(subset(Epa1, padj<0.1 & abs(log2FoldChange)>1.5), points(log2FoldChange, -log10(padj), pch=20, col="red", cex=0.5))
with(subset(Epa1, padj<0.1 & abs(log2FoldChange)>1.5), text(log2FoldChange+0.05, -log10(padj)+0.1, Gene, pch=20, col="red", cex=0.75))

heatmap.2(assay(vstEp)[ Epa$Gene, ], scale="row", trace="none", ColSideColors = as.character(as.numeric(vstEp$Growth)), col=RdBu[11:1], main="Ep genes")

write.csv(Epa1, file="nature-tables/Supp_2_DEG_Ep_growing_stable.csv")
```


