# Cancer pathways

## ssGSEA

Figure out the signalling pathway expression (e.g using ssGSEA)

```{r}
gsList=list()
for (i in 1:length(PathwayListRat)){
gsList[[i]]=GeneSet(na.omit(PathwayListRat[[i]]), setName = names(PathwayListRat)[i], geneIdType=SymbolIdentifier())
}

CancerRatSig=GeneSetCollection(gsList)
# 
Cancer_counts=gsva(assay(dds)[ ,-grep("CD45", colnames(assay(dds)))], CancerRatSig, mx.diff=T, method="ssgsea", verbose=F, parallel.sz=1, kcdf="Poisson", ssgsea.norm=T)

Cancer_counts2=gsva(allTPMFinal[ ,-grep("CD45", colnames(assay(dds)))], CancerRatSig, mx.diff=T, method="ssgsea", verbose=F, parallel.sz=1, kcdf="Poisson", ssgsea.norm=T)

# doesn't seem to matter that we convert to TPM or not. Also try the TPMs from the RSEM package
#Imm_counts2=gsva(allTPMFinal,ImmRatSig, mx.diff=T, method="ssgsea", verbose=F, parallel.sz=1, kcdf="Poisson", ssgsea.norm=T)

#pdf(sprintf("rslt/signatureAnalysis/Main_cancer_Pathways_%s.pdf", Sys.Date()), width=6, height=7)
# palette()[as.numeric(infoTableFinal$Batch[match(colnames(Cancer_counts), infoTableFinal$SampleID)])],   
heatmap.plus(Cancer_counts2, col=brewer.pal(11, "RdBu")[11:1], trace="none", ColSideColors =cbind(palette()[as.numeric(infoTableFinal$Treatment[match(colnames(Cancer_counts), infoTableFinal$SampleID)])+4],                             palette()[as.numeric(infoTableFinal$Fraction[match(colnames(Cancer_counts), infoTableFinal$SampleID)])],
             palette()[as.numeric(infoTableFinal$Growth[match(colnames(Cancer_counts), infoTableFinal$SampleID)])]), scale="row", symkey=T)


CCount=Cancer_counts2[ , grep("Ep", colnames(Cancer_counts2))]

heatmap.plus(CCount, col=brewer.pal(11, "RdBu")[11:1], trace="none", ColSideColors =cbind(palette()[as.numeric(infoTableFinal$Treatment[match(colnames(CCount), infoTableFinal$SampleID)])+4],                             palette()[as.numeric(infoTableFinal$Fraction[match(colnames(CCount), infoTableFinal$SampleID)])],
             palette()[as.numeric(infoTableFinal$Growth[match(colnames(CCount), infoTableFinal$SampleID)])]), scale="row", symkey=T,
             main="Ep only samples")

CCount=Cancer_counts2[ , grep("DN", colnames(Cancer_counts2))]

heatmap.plus(CCount, col=brewer.pal(11, "RdBu")[11:1], trace="none", ColSideColors =cbind(palette()[as.numeric(infoTableFinal$Treatment[match(colnames(CCount), infoTableFinal$SampleID)])+4],                             palette()[as.numeric(infoTableFinal$Fraction[match(colnames(CCount), infoTableFinal$SampleID)])],
             palette()[as.numeric(infoTableFinal$Growth[match(colnames(CCount), infoTableFinal$SampleID)])]), scale="row", symkey=T,
             main="DN only samples")

cancerCountsMelt=melt(Cancer_counts)
cancerCountsMelt$Fraction=infoTableFinal$Fraction[match(cancerCountsMelt$Var2, infoTableFinal$SampleID)]
cancerCountsMelt$Treatment=infoTableFinal$Treatment[match(cancerCountsMelt$Var2, infoTableFinal$SampleID)]

ggplot(data=cancerCountsMelt, aes(x=Treatment, y=value, col=Fraction))+geom_boxplot()+facet_wrap(~Var1, scales="free_y")


#dev.off()

```


### regression model for covariates

Linear model: Create a linear model to explain the above:

Score ~ Fraction + Treatment + CD8Fraction + Spatial Score + GrowthRate (+Tumor Size?) + PAM50

```{r, eval=F}
mx1=match(colnames(Cancer_counts2), infoTableFinal$SampleID)

infoTabTemp=infoTableFinal[mx1, ]
infoTabTemp$Fraction=factor(infoTabTemp$Fraction)
levels(infoTabTemp$Treatment)=c("Vehicle", "LY", "PDL1", "PDL1+LY")
#infoTabTemp$PAM50_v2="Basal"
#infoTabTemp$PAM50_v2[grep("Lum", infoTabTemp$PAM50)]="Lum"

## First do a score based on DN vs Ep

a1=glm(infoTableFinal$Fraction[mx1]~0+t(Cancer_counts2), family="binomial")
summary(a1)


MatrixSumm=matrix(NA, nrow=nrow(Cancer_counts2), ncol=11)
PValsSumm=MatrixSumm


for (i in 1:nrow(Cancer_counts2)){
  a1=glm(Cancer_counts2[i, ]~Fraction+Treatment+CD8Frac+MHcut+GrowthRate+TumSize+SpatialManual+PAM50_v2, data=infoTabTemp)
  MatrixSumm[i, ]=coefficients(a1)
  PValsSumm[i, ]=summary(a1)$coefficients[ ,4]
}

colnames(PValsSumm)=names(coefficients(a1))
rownames(PValsSumm)=rownames(Cancer_counts2)
colnames(MatrixSumm)=colnames(PValsSumm)
rownames(MatrixSumm)=rownames(PValsSumm)

MatrixSumm2=MatrixSumm
MatrixSumm2[which(PValsSumm>0.1, arr.ind = T)]=0

#pdf(sprintf("rslt/signatureAnalysis/cancer_pathway_variables_%s.pdf", Sys.Date()), height=6, width=6)
par(oma=c(4, 0,0,0))

heatmap.2(MatrixSumm2[ ,-1], col=RdBu[11:1], trace="none", main="Contributors to signalling pathway: All variables")

MatrixSumm=matrix(NA, nrow=nrow(Cancer_counts2), ncol=9)
PValsSumm=MatrixSumm

for (i in 1:nrow(Cancer_counts2)){
  a1=glm(Cancer_counts2[i, infoTabTemp$Fraction=="DN"]~Treatment+CD8Frac+MHcut+GrowthRate+TumSize+PAM50_v2, data=infoTabTemp[infoTabTemp$Fraction=="DN", ])
  MatrixSumm[i, ]=coefficients(a1)
  PValsSumm[i, ]=summary(a1)$coefficients[ ,4]
}

colnames(PValsSumm)=names(coefficients(a1))
rownames(PValsSumm)=rownames(Cancer_counts2)
colnames(MatrixSumm)=colnames(PValsSumm)
rownames(MatrixSumm)=rownames(PValsSumm)

MatrixSumm2=MatrixSumm
MatrixSumm2[which(PValsSumm>0.1, arr.ind = T)]=0

heatmap.2(MatrixSumm2[ ,-1], col=RdBu[11:1], trace="none", main="Contributors to signalling pathway: DN")

MatrixSumm=matrix(NA, nrow=nrow(Cancer_counts2), ncol=9)
PValsSumm=MatrixSumm

for (i in 1:nrow(Cancer_counts2)){
  a1=glm(Cancer_counts2[i, infoTabTemp$Fraction=="Ep"]~Treatment+CD8Frac+MHcut+GrowthRate+TumSize+PAM50_v2, data=infoTabTemp[infoTabTemp$Fraction=="Ep", ])
  MatrixSumm[i, ]=coefficients(a1)
  PValsSumm[i, ]=summary(a1)$coefficients[ ,4]
}

colnames(PValsSumm)=names(coefficients(a1))
rownames(PValsSumm)=rownames(Cancer_counts2)
colnames(MatrixSumm)=colnames(PValsSumm)
rownames(MatrixSumm)=rownames(PValsSumm)

MatrixSumm2=MatrixSumm
MatrixSumm2[which(PValsSumm>0.1, arr.ind = T)]=0

heatmap.2(MatrixSumm2[ ,-1], col=RdBu[11:1], trace="none", main="Contributors to signalling pathway: Ep")

#dev.off()
```

Associations: 
* notch signalling with LY treatment
* TGFB with PDL1+LY treatment
* Ras and Wnt signalling lower CD8Fraction
* TP53 with CD8 fraction

### TSG vs Oncogenes

Do the same scoring, taking into account oncogenic vs tumor suppressor activity?
Usually, there is no difference. The pathway is increased either way

```{r}
gsListUP=list()
for (i in 1:length(PathwayListRat)){
gsListUP[[i]]=GeneSet(na.omit(PathwayListRat[[i]][which(PathwaySign[[i]]=="OG")]), setName = paste(names(PathwayListRat)[i], "OG", sep="_"), geneIdType=SymbolIdentifier())
}
gsListDOWN=list()
for (i in 1:length(PathwayListRat)){
gsListDOWN[[i]]=GeneSet(na.omit(PathwayListRat[[i]][which(PathwaySign[[i]]=="TSG")]), setName = paste(names(PathwayListRat)[i], "TS", sep="_"), geneIdType=SymbolIdentifier())
}
AllGS=c(gsListUP, gsListDOWN)

CancerRatSig_Dir=GeneSetCollection(AllGS)
# 
Cancer_counts_dir=gsva(assay(dds)[ ,-grep("CD45", colnames(assay(dds)))], CancerRatSig_Dir, mx.diff=T, method="ssgsea", verbose=F, parallel.sz=1, kcdf="Poisson", ssgsea.norm=T)
# doesn't seem to matter that we convert to TPM or not. Also try the TPMs from the RSEM package
#Imm_counts2=gsva(allTPMFinal,ImmRatSig, mx.diff=T, method="ssgsea", verbose=F, parallel.sz=1, kcdf="Poisson", ssgsea.norm=T)

#pdf(sprintf("rslt/signatureAnalysis/Main_cancer_Pathways_w_direction_%s.pdf", Sys.Date()), width=6, height=7)


heatmap.plus(Cancer_counts_dir, col=brewer.pal(11, "RdBu")[11:1], trace="none", ColSideColors =cbind(palette()[as.numeric(infoTableFinal$Treatment[match(colnames(Cancer_counts), infoTableFinal$SampleID)])+4], palette()[as.numeric(infoTableFinal$Batch[match(colnames(Cancer_counts), infoTableFinal$SampleID)])], palette()[as.numeric(infoTableFinal$Fraction[match(colnames(Cancer_counts), infoTableFinal$SampleID)])]), scale="row")

cancerCountsMelt_dir=melt(Cancer_counts_dir)
cancerCountsMelt_dir$Var1=as.character(cancerCountsMelt_dir$Var1)
cancerCountsMelt_dir$Fraction=infoTableFinal$Fraction[match(cancerCountsMelt_dir$Var2, infoTableFinal$SampleID)]
cancerCountsMelt_dir$Treatment=infoTableFinal$Treatment[match(cancerCountsMelt_dir$Var2, infoTableFinal$SampleID)]
cancerCountsMelt_dir$Dir=substr(cancerCountsMelt_dir$Var1,  nchar(cancerCountsMelt_dir$Var1)-1, nchar(cancerCountsMelt_dir$Var1))

ggplot(data=cancerCountsMelt_dir, aes(x=Treatment, y=value, col=Fraction))+geom_boxplot()+facet_wrap(~Var1, scales="free_y")

#dev.off()

```

Do a regression analysis for TSG vs ONCO signatures in this pathway


```{r, eval=F}
mx1=match(colnames(Cancer_counts_dir), infoTableFinal$SampleID)

infoTabTemp=infoTableFinal[mx1, ]
infoTabTemp$Fraction=factor(infoTabTemp$Fraction)
levels(infoTabTemp$Treatment)=c("Vehicle", "LY", "PDL1", "PDL1+LY")

## First do a score based on DN vs Ep

a1=glm(infoTableFinal$Fraction[mx1]~0+t(Cancer_counts_dir), family="binomial")
summary(a1)


MatrixSumm=matrix(NA, nrow=nrow(Cancer_counts_dir), ncol=10)
PValsSumm=MatrixSumm


for (i in 1:nrow(Cancer_counts_dir)){
  a1=glm(Cancer_counts_dir[i, ]~Fraction+Treatment+CD8Frac+MHcut+GrowthRate+TumSize+SpatialManual, data=infoTabTemp)
  MatrixSumm[i, ]=coefficients(a1)
  PValsSumm[i, ]=summary(a1)$coefficients[ ,4]
}

colnames(PValsSumm)=names(coefficients(a1))
rownames(PValsSumm)=rownames(Cancer_counts_dir)
colnames(MatrixSumm)=colnames(PValsSumm)
rownames(MatrixSumm)=rownames(PValsSumm)

MatrixSumm2=MatrixSumm
MatrixSumm2[which(PValsSumm>0.1, arr.ind = T)]=0

#pdf(sprintf("rslt/signatureAnalysis/cancer_pathway_variables_withdirection_%s.pdf", Sys.Date()), height=6, width=6)
par(oma=c(4, 0,0,0))

heatmap.2(MatrixSumm2[ ,-1], col=RdBu[11:1], trace="none", main="Contributors to signalling pathway: All variables")

MatrixSumm=matrix(NA, nrow=nrow(Cancer_counts_dir), ncol=8)
PValsSumm=MatrixSumm

for (i in 1:nrow(Cancer_counts_dir)){
  a1=glm(Cancer_counts_dir[i, infoTabTemp$Fraction=="DN"]~Treatment+CD8Frac+MHcut+GrowthRate+TumSize, data=infoTabTemp[infoTabTemp$Fraction=="DN", ])
  MatrixSumm[i, ]=coefficients(a1)
  PValsSumm[i, ]=summary(a1)$coefficients[ ,4]
}

colnames(PValsSumm)=names(coefficients(a1))
rownames(PValsSumm)=rownames(Cancer_counts_dir)
colnames(MatrixSumm)=colnames(PValsSumm)
rownames(MatrixSumm)=rownames(PValsSumm)

MatrixSumm2=MatrixSumm
MatrixSumm2[which(PValsSumm>0.1, arr.ind = T)]=0

heatmap.2(MatrixSumm2[ ,-1], col=RdBu[11:1], trace="none", main="Contributors to signalling pathway: DN")

MatrixSumm=matrix(NA, nrow=nrow(Cancer_counts_dir), ncol=8)
PValsSumm=MatrixSumm

for (i in 1:nrow(Cancer_counts_dir)){
  a1=glm(Cancer_counts_dir[i, infoTabTemp$Fraction=="Ep"]~Treatment+CD8Frac+MHcut+GrowthRate+TumSize, data=infoTabTemp[infoTabTemp$Fraction=="Ep", ])
  MatrixSumm[i, ]=coefficients(a1)
  PValsSumm[i, ]=summary(a1)$coefficients[ ,4]
}

colnames(PValsSumm)=names(coefficients(a1))
rownames(PValsSumm)=rownames(Cancer_counts_dir)
colnames(MatrixSumm)=colnames(PValsSumm)
rownames(MatrixSumm)=rownames(PValsSumm)

MatrixSumm2=MatrixSumm
MatrixSumm2[which(PValsSumm>0.1, arr.ind = T)]=0

heatmap.2(MatrixSumm2[ ,-1], col=RdBu[11:1], trace="none", main="Contributors to signalling pathway: Ep")

#dev.off()

```

Looking at up vs down genes: are they differentially expressed?

```{r, eval=F}
AllSearch=lapply(AllGS, function(x) geneIds(x))
AllSearchUn=unlist(AllSearch)

DiffGenes=AllSearchUn[which(AllSearchUn%in%rownames(resDNeprslt2))]
TypeGenes=sapply(AllSearch, function(x) which(DiffGenes%in%x))

heatmap.2(assay(vsd)[ match(DiffGenes, rownames(assay(vsd))), -grep("CD45", colnames(assay(vsd)))],
          trace="none", col=RdBu[11:1], ColSideColors = palette()[colSide], scale="row")
```

