# TCGA: Associate Epcam+ inflammatory with survival

Associate the signature with outcome in TCGA: Load in TCGA right now

```{r}
TCGArsem=read.delim("../data/TCGA/BRCA.rnaseqv2__illuminahiseq_rnaseqv2__unc_edu__Level_3__RSEM_genes_normalized__data.data.txt", sep="\t")
rownames(TCGArsem)=TCGArsem[ ,1]
TCGArsem=TCGArsem[, grep("01A", colnames(TCGArsem))]
colnames(TCGArsem)=substr(colnames(TCGArsem), 1, 12)
colnames(TCGArsem)=gsub("\\.", "-", colnames(TCGArsem))
TCGArsem=TCGArsem[-1, ]

TCGArsem2=apply(TCGArsem, 2, as.numeric)
#TCGArsem2=t(TCGArsem2)
#colnames(TCGArsem2)=colnames(TCGArsem)
#TCGArsem2=data.frame(TCGArsem2)
rownames(TCGArsem2)=rownames(TCGArsem)
rownames(TCGArsem2)=sapply(strsplit(rownames(TCGArsem2), "\\|"), function(x) x[1])
TCGArsem2=TCGArsem2[-which(rownames(TCGArsem2)=="?"), ]

load("../data/TCGA/BrClin_clinical_Nov2017.RData")

ax1=match(BrClin$Patient.ID, colnames(TCGArsem))

BrClin=BrClin[-which(is.na(ax1)), ]
TCGArsem=TCGArsem[ , na.omit(ax1)]

## load in the inflammation subtype information
ThorssData=read.xlsx("../data/TCGA/Thorsson2018_table1.xlsx",1)
```


Edit the clinical data to make sure data is censored at 60 months (5 years) and that stage is given an integer value (no 2A, 2b etc.) for easier comparisons

```{r}
#m1=match(colnames(TCGAssgsea), BrClin$Patient.ID)
BrClin$Stage=(substr(BrClin$American.Joint.Committee.on.Cancer.Tumor.Stage.Code, 2, 2))
BrClin$Stage[which(BrClin$Stage==""|BrClin$Stage=="X")]=NA

BrClin$Overall.Survival..Months.[which(BrClin$Overall.Survival..Months.>=60)]=60
BrClin$Overall.Survival.Status[which(BrClin$Overall.Survival..Months.>=60)]="LIVING"

BrClin$Disease.Free..Months.[which(BrClin$Disease.Free..Months.>=60)]=60
BrClin$Disease.Free.Status[which(BrClin$Disease.Free..Months.>=60)]="Disease Free"

```


```{r, eval=F}
## Load the old file
Nickold=read.csv("outputs/EpCAM_IFNg_low_growing_vs_stable.filtered.csv", row.names = 1)

GlistGrowing=rownames(Nickold)[which(sign(Nickold$log2FoldChange)<0 & Nickold$baseMean>100)]
Glistnon=rownames(Nickold)[which(sign(Nickold$log2FoldChange)>0 & Nickold$baseMean>100)]

#alist=results(vstLumRes)
```

## Associating CD74 with phenotype and outcome

The Thorsson data has pre-calculated scores for:

* immune subtypes
* leukocyte fractins
* proportion of data with coding mutations
* TCR shannon index

Z-Scale CD74 scores prior to analysis:

```{r 5d, fig.cap="CD74 assoc with patient data"}
TCGAssgsea=scale(TCGArsem2[match("CD74", rownames(TCGArsem2)), ])
m1=match(colnames(TCGArsem2), BrClin$Patient.ID)

tinfo=data.frame(pam=BrClin$PAM50[m1], OSM=BrClin$Overall.Survival..Months.[m1],
                 OSS=BrClin$Overall.Survival.Status[m1], cd74=TCGAssgsea,
                 DFS=BrClin$Disease.Free.Status[m1],
                 DFM=BrClin$Disease.Free..Months.[m1],
                 Stage=BrClin$Stage[m1])

ThorssData=ThorssData[which(ThorssData$TCGA.Study=="BRCA"), ]

n1=match(rownames(tinfo), ThorssData$TCGA.Participant.Barcode)
tinfo$immSub=ThorssData$Immune.Subtype[n1]
tinfo$leukFrac=as.numeric(ThorssData$Leukocyte.Fraction[n1])
tinfo$codingMut=as.numeric(ThorssData$Nonsilent.Mutation.Rate[n1])
tinfo$TCRshann=as.numeric(ThorssData$TCR.Shannon[n1])


#pdf("figure-outputs/Figure5_CD74_output.pdf", height=7, width=9)

par(mfrow=c(2,3))
boxplot(tinfo$cd74~tinfo$pam, ylab="gene exp", main="PAM50")
boxplot(tinfo$cd74~tinfo$immSub, ylab="gene exp", main="Immune sbtype")
ax=cor.test(tinfo$cd74, log10(tinfo$codingMut+1), method="pearson")
smoothScatter(tinfo$cd74, log10(tinfo$codingMut+1), xlab="gene exp", ylab="log10 mut")
text(4,0.5, sprintf("cor=%s, p=%s", round(ax$estimate, 2), round(ax$p.value, 2)))
ax=cor.test(tinfo$cd74, tinfo$leukFrac, method="pearson")
smoothScatter(tinfo$cd74, (tinfo$leukFrac), xlab="gene exp", ylab="leuk frac")
text(4,0.2, sprintf("cor=%s, p=%s", round(ax$estimate, 2), round(ax$p.value, 2)))
ax=cor.test(tinfo$cd74, tinfo$TCRshann, method="pearson")
smoothScatter(tinfo$cd74, tinfo$TCRshann, xlab="gene exp", ylab="TCR diversity")
text(4,1, sprintf("cor=%s, p=%s", round(ax$estimate, 2), round(ax$p.value, 2)))
hist(tinfo$cd74)

#dev.off()

write.csv(tinfo, file="nature-tables/5d-TCGA_data.csv")
```

We can also check if there is an association with survival:

```{r}

axD=Surv(tinfo$DFM, ifelse(tinfo$DFS=="Recurred/Progressed", 1, ifelse(tinfo$DFS=="DiseaseFree", 0, NA)))

axO=Surv(tinfo$OSM, ifelse(tinfo$OSS=="DECEASED", 1, ifelse(tinfo$OSS=="LIVING", 0, NA)))

SurvOSS=coxph(axO~tinfo$pam+tinfo$Stage+tinfo$cd74)
SurvDFS=coxph(axD~tinfo$pam+tinfo$Stage+tinfo$cd74)

ao=summary(SurvOSS)
ao
as=summary(SurvDFS)
as
```
Note that CD74 is associated with survival: The harzards ratio is `r ao$coefficients[5,2]` (95CI:`r ao$conf.int[ 5,3:4]`), p value `r ao$logtest[3]`. 

This means that this is associated with survival regardless of subtype. We can look at these values in forestplots

```{r 5e, fig.cap="forest plot CD74"}

data1=data.frame(X1=c("","PAM50", "", "", "", "", "Stage", "","","", "CD74"),
                X2=c("","Basal", "Her2", "LumA", "LumB", "Normal" ,"1", "2", "3", "4", ""),
                X3=c("NSamp",  table(tinfo$pam), table(tinfo$Stage), length(na.omit(tinfo$cd74))),
                X4=c("HR", "", round(summary(SurvOSS)$coefficients[ 1:4,2],2), "",
                     round(summary(SurvOSS)$coefficients[ 5:8,2],2)))

mdata=data.frame(summary(SurvOSS)$conf.int[ ,c(1,3:4)])
mdata=rbind(NA,NA, mdata[1:4, ], NA, mdata[5:8, ])

#pdf("figure-outputs/Figure5_CD74_forestplot.pdf", width=5, height=5)

forestplot(data1, mdata, xlog=T,
           boxsize=0.5)

#dev.off()

write.csv(cbind(data1, mdata), file="nature-tables/5e_forest_plot.csv")

```





We can look below what the association with subtype is in KM curves:

```{r Ext5e-os, fig.cap="OS CD74 by subtype", fig.height=8}
par(mfrow=c(2,2))

Xind=c("Basal", "Her2", "LumA", "LumB")

for (i in Xind){
  l1=which(tinfo$pam==i)
  ax=Surv(tinfo$OSM[l1], ifelse(tinfo$OSS[l1]=="DECEASED", 1, ifelse(tinfo$OSS[l1]=="LIVING", 0, NA)))
  TCGAvalCut=cut(tinfo$cd74[l1], quantile(tinfo$cd74[l1], c(0, 0.33, 0.67, 1)), c("L","M", "P"))
  ax1=summary(coxph(ax~TCGAvalCut+tinfo$Stage[l1]))
  ax2=plot(survfit(ax~TCGAvalCut), main=paste(i, "CD74 gene exp"), col=brewer.pal(3, "Blues"),lwd=2, xlab="Time (months)", ylab="OS", mark.time=T)
  text(20, 0.2, sprintf("HR:%s (%s-%s) p=%s", round(ax1$coefficients[2,2 ],2), 
                     round(ax1$conf.int[2,3 ],2), round(ax1$conf.int[2, 4],2),round(ax1$logtest[3],2)))
}
```


```{r Ext5e-dfs, fig.cap="DFS CD74 by subtype"}

par(mfrow=c(2,2))

for (i in Xind){
   l1=which(tinfo$pam==i)
  ax=Surv(tinfo$DFM[l1], ifelse(tinfo$DFS[l1]=="Recurred/Progressed", 1, ifelse(tinfo$DFS[l1]=="DiseaseFree", 0, NA)))
  TCGAvalCut=cut(tinfo$cd74[l1], quantile(tinfo$cd74[l1], c(0, 0.33, 0.67, 1)), c("L","M", "P"))
  ax1=summary(coxph(ax~TCGAvalCut+tinfo$Stage[l1]))
  ax2=plot(survfit(ax~TCGAvalCut), main=paste(i, "CD74 gene exp"), col=brewer.pal(3, "Blues"),lwd=2, xlab="Time (months)", ylab="DFS", mark.time=T)
  text(20, 0.2, sprintf("HR:%s (%s-%s) p=%s", round(ax1$coefficients[2,2 ],2), 
                     round(ax1$conf.int[2,3 ],2), round(ax1$conf.int[2, 4],2),round(ax1$logtest[3],2)))
}
```


## Signature: Lum cases non-inflammatory: growing vs stable

Here, use the signature which was applied before to see the difference between the non-inflammatory growing vs stable samples

```{r, results='hide'}
GlistGrowing=rownames(vstLumRes)[which(vstLumRes$log2FoldChange>1.5 & vstLumRes$baseMean>100 & vstLumRes$padj<0.05 )]
Glistnon=rownames(vstLumRes)[which(vstLumRes$log2FoldChange<(-1.5) & vstLumRes$baseMean>100 & vstLumRes$padj<0.05 )]

HumGeneList=SymHum2Rat$HGNC.symbol[match(GlistGrowing, SymHum2Rat$RGD.symbol)]
HumGeneList2=Rat2Hum$HGNC.symbol[match(GlistGrowing, Rat2Hum$RGD.symbol)]

TCGAssgsea=gsva((TCGArsem2), list(grow=na.omit(HumGeneList2), 
                                                stab=Rat2Hum$HGNC.symbol[match(Glistnon, Rat2Hum$RGD.symbol)]), method="ssgsea", ssgsea.norm=T)

par(mfrow=c(1,2))
hist(TCGAssgsea[1, ], main="stable")
hist(TCGAssgsea[2, ], main="growing")
plot(TCGAssgsea[1, ], TCGAssgsea[2, ], xlab="stable", ylab="growing")
```

Below are the survival curves:

```{r, fig.height=8}
#pdf("figure-outputs/Fig5_survival_analysis_Lum_samples_baseMean_greater100.pdf", height=8, width = 8)

par(mfrow=c(2,2))

Xind=c("Basal", "Her2", "LumA", "LumB")

for (i in Xind){
  l1=which(BrClin$PAM50==i)
  lx2=match(BrClin$Patient.ID[l1], colnames(TCGAssgsea))
  ax=Surv(BrClin$Overall.Survival..Months.[l1], ifelse(BrClin$Overall.Survival.Status[l1]=="DECEASED", 1, ifelse(BrClin$Overall.Survival.Status[l1]=="LIVING", 0, NA)))
   axb=coxph(ax~TCGAssgsea[1, lx2]+BrClin$Stage[l1])
  axc=summary(axb)
  TCGAvalCut=cut(TCGAssgsea[1, lx2], quantile(TCGAssgsea[1, lx2], c(0, 0.33, 0.67, 1)), c("L","M", "P"))
  ax2=plot(survfit(ax~TCGAvalCut), main=paste(i, "lum stable"), col=brewer.pal(3, "Blues"),lwd=2, xlab="Time (months)", ylab="OS", mark.time=T)
   text(10, 0, sprintf("univ. cts. var. HR=%s (%s-%s), p=%s", round(axc$coefficients[1,2], 2),
                      round(axc$conf.int[1,3],2), round(axc$conf.int[1,4],2),round(axc$logtest[3],2 )))
  
}


for (i in Xind){
  l1=which(BrClin$PAM50==i)
  lx2=match(BrClin$Patient.ID[l1], colnames(TCGAssgsea))
  ax=Surv(BrClin$Disease.Free..Months.[l1], ifelse(BrClin$Disease.Free.Status[l1]=="Recurred/Progressed", 1, ifelse(BrClin$Disease.Free.Status[l1]=="DiseaseFree", 0, NA)))
  axb=coxph(ax~TCGAssgsea[1, lx2]+BrClin$Stage[l1])
  axc=summary(axb)
  
  TCGAvalCut=cut(TCGAssgsea[1, lx2], quantile(TCGAssgsea[1, lx2], c(0, 0.33, 0.67, 1)), c("L","M", "P"))
  ax2=plot(survfit(ax~TCGAvalCut), main=paste(i, "lum stable"), col=brewer.pal(3, "Blues"),lwd=2, xlab="Time (months)", ylab="DFS", mark.time=T)
  text(10, 0, sprintf("univ. cts. var. HR=%s (%s-%s), p=%s", round(axc$coefficients[1,2], 2),
                      round(axc$conf.int[1,3],2), round(axc$conf.int[1,4],2),round(axc$logtest[3],2 )))
}
```

Below are firstly the OS curves for the growing signature

```{r 5i, fig.cap="OS: growing signature", fig.height=8}
par(mfrow=c(2,2))

Xind=c("Basal", "Her2", "LumA", "LumB")

for (i in Xind){
  l1=which(BrClin$PAM50==i)
   lx2=match(BrClin$Patient.ID[l1], colnames(TCGAssgsea))
  ax=Surv(BrClin$Overall.Survival..Months.[l1], ifelse(BrClin$Overall.Survival.Status[l1]=="DECEASED", 1, ifelse(BrClin$Overall.Survival.Status[l1]=="LIVING", 0, NA)))
   axb=coxph(ax~TCGAssgsea[2, lx2]+BrClin$Stage[l1])
  axc=summary(axb)
  
  TCGAvalCut=cut(TCGAssgsea[2, lx2], quantile(TCGAssgsea[2, lx2], c(0, 0.33, 0.67, 1)), c("L","M", "P"))
  ax2=plot(survfit(ax~TCGAvalCut), main=paste(i, "lum growing"), col=brewer.pal(3, "Blues"),lwd=2, xlab="Time", ylab="OS", mark.time=T)
   text(10, 0, sprintf("univ. cts. var. HR=%s (%s-%s), p=%s", round(axc$coefficients[1,2], 2),
                      round(axc$conf.int[1,3],2), round(axc$conf.int[1,4],2),round(axc$logtest[3],2 )))
  
}
```

Followed by the DFS plots for this signature

```{r 5i-dfs, fig.cap="DFS forgrowing signature", fig.height=8}
par(mfrow=c(2,2))

for (i in Xind){
  l1=which(BrClin$PAM50==i)
   lx2=match(BrClin$Patient.ID[l1], colnames(TCGAssgsea))
  ax=Surv(BrClin$Disease.Free..Months.[l1], ifelse(BrClin$Disease.Free.Status[l1]=="Recurred/Progressed", 1, ifelse(BrClin$Disease.Free.Status[l1]=="DiseaseFree", 0, NA)))
   axb=coxph(ax~TCGAssgsea[2, lx2]+BrClin$Stage[l1])
  axc=summary(axb)
  
  TCGAvalCut=cut(TCGAssgsea[2, lx2], quantile(TCGAssgsea[2, lx2], c(0, 0.33, 0.67, 1)), c("L","M", "P"))
  ax2=plot(survfit(ax~TCGAvalCut), main=paste(i, "lum growing"), col=brewer.pal(3, "Blues"),lwd=2, xlab="Time (months)", ylab="DFS", mark.time=T)
    text(10, 0, sprintf("univ. cts. var. HR=%s (%s-%s), p=%s", round(axc$coefficients[1,2], 2),
                      round(axc$conf.int[1,3],2), round(axc$conf.int[1,4],2),round(axc$logtest[3],2 )))
  
}
```

The gene lists are 

Stable:

`r GlistGrowing` 

Growing:

`r Glistnon`


We can see whether this signature associates with leukocyte content and TCR diversity in the TCGA cohort:

```{r 5h, fig.cap="luminal signature associated with clinical variables"}

l1=which(BrClin$PAM50=="LumA")
lx2=match(BrClin$Patient.ID[l1], colnames(TCGAssgsea))

lx3=match(BrClin$Patient.ID[l1], ThorssData$TCGA.Participant.Barcode)

newinfoTab=data.frame(pat=BrClin$Patient.ID[l1], TCGAval=TCGAssgsea[2, lx2],
                      leuk=as.numeric(ThorssData$Leukocyte.Fraction[lx3]),
                      TCRshann=as.numeric(ThorssData$TCR.Shannon[lx3]),
                      immSub=ThorssData$Immune.Subtype[lx3])


#pdf("figure-outputs/Fig5_assoc_growing_stable_signature.pdf", height=6, width=9)
par(mfrow=c(1,2))
ax=cor.test(newinfoTab$TCGA, newinfoTab$leuk, method="pearson")
smoothScatter(newinfoTab$TCGA, newinfoTab$leuk)
text(0.6,0.6, sprintf("cor=%s, p=%s", round(ax$estimate, 2), round(ax$p.value, 2)))
ax=cor.test(newinfoTab$TCGA, newinfoTab$TCRshann, method="pearson")
smoothScatter(newinfoTab$TCGA, newinfoTab$TCRshann)
text(0.6,5, sprintf("cor=%s, p=%s", round(ax$estimate, 2), round(ax$p.value, 2)))
```

As well as whether it associates with any of the previously identified immune subtypes:

```{r thorrs-subtypes}
ggplot(newinfoTab, aes(x=immSub, y=TCGAval))+geom_boxplot()+geom_point()+theme_bw()
write.csv(newinfoTab, file="nature-tables/5h_lum_growing_vs_stable.csv")
```

### TCGA: LumA high vs low signatures

We can pull out the samples here which have high vs low signature (in the lumA group) and perform differential gene expression analysis on these cases:

```{r TCGA-deseq}
newinfoTab$TCGAvalcut=cut(newinfoTab$TCGAval,  quantile(newinfoTab$TCGAval, c(0, 0.33, 0.67, 1) ), c("L", "M", "O"))

colDataTCGA=newinfoTab[which(newinfoTab$TCGAvalcut!="M"), ]

head(colDataTCGA)

TCGAdeseq=DESeqDataSetFromMatrix(round(TCGArsem2[ ,match(colDataTCGA$pat, colnames(TCGArsem2))]), colDataTCGA, design=~TCGAvalcut)

TCGAdeseq=DESeq(TCGAdeseq)

TCGAdeseqRes=results(TCGAdeseq, contrast=c("TCGAvalcut", "L", "O"))

write.csv(TCGAdeseqRes, file="nature-tables/5J_TCGA.csv")

TCGAdeseqRes2=TCGAdeseqRes[which(TCGAdeseqRes$padj<0.05 & abs(TCGAdeseqRes$log2FoldChange)>1.5 
                                 & TCGAdeseqRes$baseMean>100), ]
```

We can plot the differential genes here in a volcano plot:

```{r tcga-volcano, fig.cap="yet another volcano plot"}
#pdf("figure-outputs/Fig5J_TCGA_deg.pdf", height=8, width=8)
with(TCGAdeseqRes, plot(log2FoldChange, -log10(padj), pch=20, main="Volcano plot: Low exp (+) vs high (-)", cex=1.0, xlab=bquote(~Log[2]~fold~change), ylab=bquote(~-log[10]~Q~value)))
with(subset(TCGAdeseqRes, padj<0.1 & abs(log2FoldChange)>1.5), points(log2FoldChange, -log10(padj), pch=20, col="red", cex=0.5))
with(subset(TCGAdeseqRes, padj<0.1 & abs(log2FoldChange)>1.5), text(log2FoldChange+0.05, -log10(padj)+0.05, rownames(TCGAdeseqRes2), pch=20, col="red", cex=0.75))
#dev.off()
```


```{r heatmap-tcga, fig.cap="because someone will ask for it"}
vstTCGA=vst(assay(TCGAdeseq))

ColSideCols=c("skyblue", "blue")[factor(TCGAdeseq$TCGAvalcut)]

ax1=vstTCGA[ match(rownames(TCGAdeseqRes2), rownames(vstTCGA)),  ]

#pdf("figure-outputs/Figure5J_XX_heatmap_LumA_TCGA.pdf", height=15, width=10)

heatmap.2(vstTCGA[ rownames(TCGAdeseqRes2),  ], scale="row", trace="none", ColSideColors = ColSideCols, col=RdBu[11:1], main="LumA TCGA")
```


also check the directions by looking at CXCl13, which is an overexpressed gene:

```{r}
boxplot(vstTCGA["CXCL13", ]~TCGAdeseq$TCGAvalcut)

```

Run gsea here

```{r TCGA-gsea, fig.cap="pointless TCGA hairball"}

hits=rownames(TCGAdeseqRes2)
fcTab=TCGAdeseqRes$log2FoldChange

names(fcTab)=rownames(TCGAdeseq)
  gscaTCGA=GSCA(listOfGeneSetCollections=ListGSC,geneList=fcTab, hits = hits)
   gscaTCGA<- preprocess( gscaTCGA, species="Hs", initialIDs="SYMBOL",
                    keepMultipleMappings=TRUE, duplicateRemoverMethod="max",
                    orderAbsValue=FALSE)
   gscaTCGA <- analyze( gscaTCGA, 
                 para=list(pValueCutoff=0.05, pAdjustMethod="BH",
                           nPermutations=100, minGeneSetSize=5,
                           exponent=1), 
                           doGSOA = F)
   
save( gscaTCGA, file="figure-outputs/TCGA_hairball.RData")
   
TermsA=sapply(strsplit(rownames(gscaTCGA@result$GSEA.results$ProcessNetworks), "_"), function(x) x[2])
TermsA[which(is.na(TermsA))]=substr(rownames(gscaTCGA@result$GSEA.results$ProcessNetworks)[which(is.na(TermsA))], 2, 50)

## check whether this runs:
gscaTCGA@result$GSEA.results$ProcessNetworks$Gene.Set.Term=TermsA
viewEnrichMap(gscaTCGA, gscs=c("ProcessNetworks"),
              allSig = TRUE, gsNameType="term" )

write.csv(gscaTCGA@result$GSEA.results$ProcessNetworks, file="nature-tables/5k_TCGA_lumA_GSEA.csv")
```

