# Trichrome staining

Quantification of trichrome staining was performed in Qupath using the following steps:

* image is loaded as a "DAB" image
* using "gold standard" trichrome-stained images with good stroma and epithelial content, estimate the stain vectors
* Perform color deconvolution
* A pixel classifier was used to estimate trichrome content
* A pixel classifier was used to estimate tumor content

A snippet of the qupath script is shown below:

```{r, eval=F, echo=T}
# insert some text here

setImageType('BRIGHTFIELD_H_DAB');
setColorDeconvolutionStains('{"Name" : "trichrome", "Stain 1" : "Hematoxylin", "Values 1" : "0.71695 0.66336 0.21432 ", "Stain 2" : "DAB", "Values 2" : "0.46299 0.76212 0.45257 ", "Background" : " 255 255 255 "}');
selectAnnotations();
addPixelClassifierMeasurements("stroma_classifier_again", "stroma_classifier_again")
selectAnnotations();
addPixelClassifierMeasurements("test_tumor", "test_tumor")

def entry = getProjectEntry()
def name = entry.getImageName() + '.txt'
def path = buildFilePath(PROJECT_BASE_DIR, 'annotation results')
mkdirs(path)
path = buildFilePath(path, name)
saveAnnotationMeasurements(path)

```

## Associations with cellular fraction (wsi)

```{r, height=6, eval=T}
# load the data
TrichromeData=read.xlsx("../data/trichrome-staining-result.xls", 1)
midx=match(TrichromeData$SampleID, gsub("_", "", Cdata$TumorID))

Cdata$Trichrome=NA
Cdata$Trichrome[na.omit(midx)]=TrichromeData$Percentage.Stroma[which(!is.na(midx))]

t2=WSIvalFracs[, match(rownames(df.Spatial), colnames(WSIvalFracs))]

df.Spatial=cbind(df.Spatial, t(t2))

#head(df.Spatial)

df.Spatial$Trichrome=NA
midx=match(TrichromeData$SampleID, rownames(df.Spatial))
df.Spatial$Trichrome[na.omit(midx)]=TrichromeData$Percentage.Stroma[-which(is.na(midx))]
# plot associations

n2=c("CD8", "EpCAM", "SMA", "Unclass", "EpCAM: SMA")

#pdf("~/Desktop/richrome-association-WSI-data-Calc2.pdf", width=6, height=5)

par(mfrow=c(2,3))
for (i in n2){
   a1=cor.test(df.Spatial$Trichrome, df.Spatial[ ,match(i, colnames(df.Spatial))], use="complete")
   n1=paste(i, "cor:", round(a1$estimate,1), "p:", round(a1$p.value,2))
   plot(df.Spatial$Trichrome, df.Spatial[ ,match(i, colnames(df.Spatial))], main=n1, xlab="Trichrome", ylab=i)
}

#dev.off()
```

## Associations with CD8 content

```{r 3gpre, fig.height=6}

par(mfrow=c(1,2))

a1=wilcox.test(df.Spatial$Trichrome~df.Spatial$IF.EpCAMcut)$p.val
boxplot(df.Spatial$Trichrome~df.Spatial$IF.EpCAMcut, main=sprintf("Ep int. fraction %s", round(a1,2)))
a1=wilcox.test(df.Spatial$Trichrome~df.Spatial$CD8Fraccut)$p.val
boxplot(df.Spatial$Trichrome~df.Spatial$CD8Fraccut, main=sprintf("CD8 fraction vs Trichrome %s", round(a1,2)))
```

## Associations with growth and treatment

```{r 3g, fig.cap="Trichrome staining with treatment"}
#pdf("~/Desktop/testoutput.pdf", height=6, width=6)
pv1=sapply(c("PDL1", "LY", "PDL1+LY"), function(x) wilcox.test(df.Spatial$Trichrome[which(df.Spatial$Treatment%in%c("Vehicle", x))]~
           df.Spatial$Treatment[which(df.Spatial$Treatment%in%c("Vehicle", x))])$p.value)
names(pv1)=c("PDL1", "LY", "PDL1+LY")

ggplot(df.Spatial[ ,c("Treatment", "Trichrome")], aes(x=Treatment, y=Trichrome, col=Treatment))+geom_boxplot()+geom_jitter()+
scale_color_manual(values=ColMerge[ ,1])+theme_bw()+ggtitle(paste(paste(names(pv1), round(pv1, 2)), collapse=";"))
```

```{r 3gb, fig.cap="Trichrome with growth"}
ggplot(df.Spatial[ ,c("Trichrome", "Growth")], aes(x=Growth, y=Trichrome, col=Growth))+geom_boxplot()+geom_jitter()+
 scale_color_manual(values=c(ColSize, "black"))+theme_bw()+ggtitle(paste("growth", round(a1a, 2)))
print(p)

write.csv(df.Spatial[ ,c("Trichrome", "Growth", "Treatment")], file="nature-tables/3g_trichrome.csv")

```

## Association with hyperinflammatory status

We can also do a boxplot for trichrome status and whether it associates with hyperinflammation in our rat samples

```{r}
vstEpInf$Inflammation3
ax1=colnames(vstEpInf)
ax1

t2=match(gsub("_", "", substr( ax1, 1, nchar(ax1)-2)), rownames(df.Spatial))
t2

```

# Writing files to file

Update the metadata file:
write to file:

* updated ext data table
* spatial information summary
* star aligned table
* colData information
* collapsed information on wsi/facs

```{r}
write.xlsx(Cdata, file=sprintf("../metadata/Carlos_samples_list_master_updated_%s.xlsx", Sys.Date()))

x1=Cdata$NewID[match(rownames(df.Spatial), gsub("_","", Cdata$TumorID))]
head(x1)
write.csv(df.Spatial, file="nature-tables/summary_spatial_information.csv")

colnames(allstarFinal)=infoTableFinal$TumorIDnew[match(colnames(allstarFinal), rownames(infoTableFinal))]
infoTableFinal2=infoTableFinal[ ,c("TumorIDnew","Growth", "CD8Frac", "TumSize", "Time.NMU2Sac", 
                                   "Age.Injection","Treatment","Cohort", "SizeCat")]
write.csv(infoTableFinal2, file="nature-tables/infoTableFinal2_output.csv")
write.csv(allstarFinal, file="nature-tables/allstarFinal_output.csv")


colTestCD8=c("CD8.EpBoundingBox", "CD8.TumSize", "CD45.FACS", "CD8.FACS", "OverallCD8_Fig4c.Manual", "TIL_Fig4d.Manual", "T.cell.CD8._TIMER", "T.cell.CD8._CIBERSORT","T.cell.CD8._CIBERSORT.ABS", "T.cell.CD8._EPIC")

# colTestCD8=c("CD8.EpDomTiles","CD8.EpBoundingBox", "CD8.TumSize", "CD45.FACS", "CD8.FACS", "OverallCD8_Fig4c.Manual", "TIL_Fig4d.Manual", )
indx1=c("CD8.WSI", "CD8Frac.WSI", "CD8_EPorSMARatio.WSI","CD8_AnyEPRatio.WSI", "CD8_EPRatio.WSI","CD8normTumSize", "CD8.EpBoundingBox", "log2CD8_EPorSMARatio.WSI", "log2CD8_EPRatio.WSI")

# df2=data.frame(CD8.WSI=t(WSIvals[ 1, ]), 
#                #CD8Frac.WSI=t(WSIvalFracs[ ,2]), 
#                CD8_EPorSMARatio.WSI=(WSIvals[1,  ]/colSums(WSIvals[2:4, ])),
#                CD8_AnyEPRatio.WSI=(WSIvals[1,  ]/colSums(WSIvals[2:3, ])),
#                CD8_EPRatio.WSI=(WSIvals[1,  ]/(WSIvals[2, ])),
#                CD8normTumSize=(WSIvals[1,  ]/df.Spatial)
#                )

```


Below is the session info

```{r}
sessionInfo()
```