# DESeq analysis

## Summary of comparisons

```{r}
#load("rslt/DESeq/all_differential_comparisons.RData")
#load("outputs/subfraction_analysis_2020-11-07.RData")
```

Quick plot of the differences in number of differential genes:

```{r}
Allchanges=rbind(UpDn1, CDUpDn1, DNUpDn1)
Allchanges=cbind(Allchanges, Frac=c("Ep", "Ep", "CD", "CD", "DN", "DN"), dir=rownames(Allchanges))

AllchangesB=melt(as.data.frame(Allchanges), id.vars = c("Frac", "dir"))

ggplot(AllchangesB, aes(x=variable, y=as.numeric(value), fill=dir))+geom_bar(stat="identity",position = "dodge")+facet_grid(Frac~., scale="free_y", space="free_y")+theme(axis.text.x = element_text(angle = 90, hjust = 1))+ylab("No. differential genes")+ggtitle("number of differential genes")+scale_y_log10()
## plot the number of samples in each comparison

Nsamp=matrix(NA, nrow=6, ncol=ncol(Allchanges))
ta=table(Epdds$Comp4)
tb=table(CDdds$Comp4)
tc=table(DNdds$Comp4)

for (i in 1:4){
  Nsamp[1:2, i]=c(ta[xsearch[i]], ta[ysearch[i]])
  Nsamp[3:4, i]=c(tb[xsearch[i]], tb[ysearch[i]])
  Nsamp[5:6, i]=c(tc[xsearch[i]], tc[ysearch[i]])
}

ta=table(Epdds$Treatment)
tb=table(CDdds$Treatment)
tc=table(DNdds$Treatment)

for (i in 1:3){
  Nsamp[1:2, i+4]=c(ta[cList[i]], ta["Vehicle"])
  Nsamp[3:4, i+4]=c(tb[cList[i]], tb["Vehicle"])
  Nsamp[5:6, i+4]=c(tc[cList[i]], tc["Vehicle"])
}

ta=table(EpddsTreatG$Comp8)
tb=table(CDddsTreatG$Comp8)
tc=table(DNddsTreatG$Comp8)

for (i in 1:4){
  Nsamp[1:2, i+7]=c(ta[xsearchA[i]], ta[ysearchB[i]])
  Nsamp[3:4, i+7]=c(tb[xsearchA[i]], tb[ysearchB[i]])
  Nsamp[5:6, i+7]=c(tc[xsearchA[i]], tc[ysearchB[i]])
}

Nsamp[ ,12]=c(table(EpddsGrowth$Growth), table(CDddsGrowth$Growth), table(DNddsGrowth$Growth))
Nsamp[ ,13]=c(table(EpddsStrMH$MHcut), table(CDddsStrMH$MHcut), table(DNddsStrMH$MHcut))
Nsamp[ ,14]=c(table(EpddsStrIF$IFcut), table(CDddsStrIF$IFcut), table(DNddsStrIF$IFcut))
Nsamp[ ,15]=c(table(EpddsStrknn$knncut), table(CDddsStrknn$knncut), table(DNddsStrknn$knncut))
Nsamp[ ,16]=c(table(Epddscd8$CD8FracCut), table(CDddscd8$CD8FracCut), table(DNddscd8$CD8FracCut))
Nsamp[ ,17]=c(table(EpddsSpatMan$SpatialManual), table(CDddsSpatMan$SpatialManual), table(DNddsSpatMan$SpatialManual))

ta=table(Epddscd8MH$cd8MH)
tb=table(CDddscd8MH$cd8MH)
tc=table(DNddscd8MH$cd8MH)


for (i in 1:4){
  Nsamp[1:2, i+17]=c(ta[xsearchC[i]], ta[ysearchD[i]])
  Nsamp[3:4, i+17]=c(tb[xsearchC[i]], tb[ysearchD[i]])
  Nsamp[5:6, i+17]=c(tc[xsearchC[i]], tc[ysearchD[i]])
}

colnames(Nsamp)=colnames(Allchanges)
Nsamp[ , 22:23]=Allchanges[ ,22:23]


Nsamp2=melt(as.data.frame(Nsamp), id.vars=c("Frac", "dir"))

ggplot(Nsamp2, aes(x=variable, y=as.numeric(value), fill=dir))+geom_bar(stat="identity",position = "dodge")+facet_grid(Frac~., scale="free_y", space="free")+theme(axis.text.x = element_text(angle = 90, hjust = 1)) + scale_fill_discrete(name = "Comparison", labels = c("numerator", "reference"))+ylab("No. samples in comparison")+ggtitle("number of samples in comparison")

#dev.off()
```

Notes:

Comparison Type | Main cell type impacted
----------------|-----------------------
Treatment | CD45+, PDL1+LY in the DN, LY may affect Ep
growth | Ep and CD45
growth in control | CD and DN
Spatial Patterns | CD45, Ep
Spatial, high cd8 | all cases

## Summary of comaprisons

Pearson correlation:

* Pearson correlation of all the samples
* Unlist all genes, and pick out genes which are impacted in at least 1 comparison, and base expression >100
* columns are colored by: Spatial Distribution (MH),  Growth, Treatment, CD8 content

Differential genes:

* top 100 variable genes
* DEG in at least 2 samples
* differential genes in each comparison shown

Gene sets:

* c2, c5, Process networks, metacore, hallmark

### EPCAM samples

Note that there are a subset of samples different from the other based on CD8, growth and MH index (6RB, 3NB, 10LD, 8LD, 11ND) but these parameters don't group the other samples well


There is a small subset of samples which have high expression of CCls and based on the top variable genes, there seems to be three types of epithelial cases. 

```{r}
SampleFrac="Ep"

allEpGenes=unlist(EpCompSig)
allEpTab=table(allEpGenes)
EpGenesList=names(allEpTab)[which(allEpTab>1)]
#EpGenesList=EpGenesList[-grep("LOC", EpGenesList)]
#EpGenesList=EpGenesList[-grep("RGD", EpGenesList)]
# also, remove all samples which have a base mean less than 10
x1=EpComp4[[1]]$baseMean[match(EpGenesList, EpComp4[[1]]$Gene)]
x2=which(x1>100)
vstEp=vst(Epdds, blind=T)

#EpddsTreatG$Comp8_v2=factor(paste(EpddsTreatG$Treatment, EpddsTreatG$Growth))

ColsideCols=hue_pal()(2)[Epdds$CD8FracCut]
ColsideColsB=hue_pal()(4)[Epdds$Treatment]
ColsideColsC=hue_pal()(2)[Epdds$Growth]
ColsideColsD=hue_pal()(2)[Epdds$MHcut]

colTable=rbind(ColsideCols, ColsideColsB, ColsideColsC, ColsideColsD)
rownames(colTable)=c("CD8", "Treatment", "growth", "MH")

#pdf(sprintf("rslt/DESeq/DEseq_comparisons_%s.pdf", SampleFrac), width=7, height=10)

sampleDist=dist(t(assay(vstEp)))
sampleDistMatrix <- as.matrix(sampleDist)
heatmap.plus(sampleDistMatrix, trace="none", scale="none", col=brewer.pal(9, "Blues")[9:1], ColSideColors = t(colTable), main=sprintf("Similarity between %s samples", SampleFrac))

# Take the top 150 variable genes 
# compute the sds for each genes
sds = rowSds(assay(vstEp))
means=rowMeans(assay(vstEp))
CVs=sds/means
CVs=CVs[-which(means<7)]
sds2=sort(CVs, decreasing = T)

heatmap.plus(assay(vstEp)[names(sds2)[1:100], ], trace="none", scale="row", col=RdBu[11:1], ColSideColors = t(colTable), main=sprintf("top 100 variable genes all %s samples", SampleFrac))


heatmap.plus(assay(vstEp)[EpGenesList[x2], ], trace="none", scale="row", col=RdBu[11:1], ColSideColors = t(colTable),
          main=sprintf("DEG at least 2 %s samples", SampleFrac))

# Make a table for these genes
FCvalues=matrix(NA, nrow=length(x2), ncol=length(EpComp4))

for (i in 1:length(EpComp4)){
  FCvalues[ ,i]=EpComp4[[i]]$log2FoldChange[match(EpGenesList[x2], EpComp4[[i]]$Gene)]
}

colnames(FCvalues)=names(EpComp4)
rownames(FCvalues)=EpGenesList[x2]


# for each list pick the top 50 diff genes FC>2 and baseman > 50
glist=NA
for (i in 1:length(EpComp4)){
  glist=c(glist, EpComp4[[i]]$Gene[which(EpComp4[[i]]$padj<0.1 & abs(EpComp4[[i]]$log2FoldChange)>2 & EpComp4[[i]]$baseMean>25)])
}

ColsideCols=SetCols[EpddsTreatG$Comp8]

glistmod=unique(glist)
glistmod=glistmod[-grep("LOC", glistmod)]
glistmod=na.omit(glistmod[-grep("RGD", glistmod)])

#heatmap.plus(assay(vstEp)[glistmod, ], trace="none", scale="row", col=RdBu[11:1], ColSideColors = t(colTable),  main=sprintf("All DEG %s samples", SampleFrac))
```

We can also view this in the form of what is highly expressed in each comparison. Below is a heatmap of the fold changes:

```{r}
# distance of the samples to one another
par(oma=c(1,0,0,0))
heatmap.2(FCvalues, trace="none", scale="row", col=RdBu[11:1], 
          main=sprintf("All comparisons %s, log2FC for each gene", SampleFrac))
#dev.off()
```

Below, we visualise the enrichment scores using different gene sets and 

```{r}
#Epismr3=ismr4

#pdf(sprintf("rslt/DESeq/GSEA_signature_comparisons_%s.pdf", SampleFrac), width=10, height=12)

for (i in 1:length(Epismr3)){
  x1=Epismr3[[i]]
  # find padj<0.1
  idx2=lapply(seq(3, ncol(x1), by=2), function(x) which(as.numeric(as.character(x1[ ,x]))<0.01))
  unList=table(unlist(idx2))
  #pickThese=names(unList)#[which(unList>1)]
  pickThese2=names(unList)[which(unList>1)]
  
  if (length(pickThese2)>100){
    temp=x1[as.numeric(pickThese2), seq(2, ncol(x1), by=2)]
    temp2=sapply(temp, function(x) as.numeric(as.character(x)))
    temp2=rowSds(abs(temp2))
    names(temp2)=pickThese2
    temp2=sort(temp2, decreasing = T)
    pickThese2=names(temp2)[1:100]
  }
  
  x23=x1[as.numeric(pickThese2), seq(2, ncol(x1), by=2)]
  
  x234=sapply(x23, function(x) as.numeric(as.character(x)))
  rownames(x234)=rownames(x23)
  
  par(oma=c(2,0,0,2))
  heatmap.2(x234, col=RdBu[11:1], scale="none", trace="none", main=sprintf("%s pathways %s fraction", 
                                                                           names(Epismr3)[i], SampleFrac))
}

#dev.off()

# boxplot(assay(vstEp)["Tap1", ]~EpddsTreatG$Treatment)
# boxplot(assay(vstEp)["Cd74", ]~EpddsTreatG$Comp4)
# boxplot(assay(vstEp)["Cd8a", ]~EpddsTreatG$Treatment)
# boxplot(assay(vstEp)["Ctla4", ]~EpddsTreatG$Treatment)
# boxplot(assay(vstEp)["Tgfbr1", ]~EpddsTreatG$Comp8)
```

### CD45 comparisons

```{r}
SampleFrac="CD"

allCDGenes=unlist(CDCompSig)
allCDTab=table(allCDGenes)
CDGenesList=names(allCDTab)[which(allCDTab>1)]
#CDGenesList=CDGenesList[-grCD("LOC", CDGenesList)]
#CDGenesList=CDGenesList[-grCD("RGD", CDGenesList)]
# also, remove all samples which have a base mean less than 10
x1=CDComp4[[1]]$baseMean[match(CDGenesList, CDComp4[[1]]$Gene)]
x2=which(x1>25)
vstCD=vst(CDdds, blind=T)

CDddsTreatG$Comp8_v2=factor(paste(CDddsTreatG$Treatment, CDddsTreatG$Growth))

ColsideCols=DiffCols[CDddsTreatG$Comp8_v2]
ColsideColsB=hue_pal()(4)[CDddsTreatG$Treatment]
ColsideColsC=hue_pal()(2)[CDddsTreatG$Growth]

colTable=rbind(ColsideCols, ColsideColsB, ColsideColsC)
rownames(colTable)=c("Treat+growth", "Treatment", "growth")



#pdf(sprintf("rslt/DESeq/DEseq_comparisons_%s.pdf", SampleFrac), width=7, height=10)

sampleDist=dist(t(assay(vstCD)))
sampleDistMatrix <- as.matrix(sampleDist)
heatmap.plus(sampleDistMatrix, trace="none", scale="none", col=brewer.pal(11, "Blues")[11:1], ColSideColors = t(colTable), main=sprintf("Similarity between %s samples", SampleFrac))

# Take the top 150 variable genes 
# compute the sds for each genes
sds = rowSds(assay(vstCD))
means=rowMeans(assay(vstCD))
CVs=sds/means
CVs=CVs[-which(means<7)]
sds2=sort(CVs, decreasing = T)

heatmap.plus(assay(vstCD)[names(sds2)[1:200], ], trace="none", scale="row", col=RdBu[11:1], ColSideColors = t(colTable), main=sprintf("top 200 variable genes all %s samples", SampleFrac))


heatmap.plus(assay(vstCD)[CDGenesList[x2], ], trace="none", scale="row", col=RdBu[11:1], ColSideColors = t(colTable),
          main=sprintf("DEG at least 2 %s samples", SampleFrac))

# Make a table for these genes
FCvalues=matrix(NA, nrow=length(x2), ncol=length(CDComp4))

for (i in 1:length(CDComp4)){
  FCvalues[ ,i]=CDComp4[[i]]$log2FoldChange[match(CDGenesList[x2], CDComp4[[i]]$Gene)]
}

colnames(FCvalues)=names(CDComp4)
rownames(FCvalues)=CDGenesList[x2]


# for each list pick the top 50 diff genes FC>2 and baseman > 50
glist=NA
for (i in 1:length(CDComp4)){
  glist=c(glist, CDComp4[[i]]$Gene[which(CDComp4[[i]]$padj<0.1 & abs(CDComp4[[i]]$log2FoldChange)>2 & CDComp4[[i]]$baseMean>25)])
}

ColsideCols=SetCols[CDddsTreatG$Comp8]

glistmod=unique(glist)
#glistmod=glistmod[-grep("LOC", glistmod)]
#glistmod=na.omit(glistmod[-grCD("RGD", glistmod)])

heatmap.plus(assay(vstCD)[na.omit(glistmod), ], trace="none", scale="row", col=RdBu[11:1], ColSideColors = t(colTable),
          main=sprintf("All DEG %s samples", SampleFrac))

# distance of the samples to one another
par(oma=c(2,0,0,0))
heatmap.2(FCvalues, trace="none", scale="row", col=RdBu[11:1], 
          main=sprintf("All comparisons %s, log2FC for each gene", SampleFrac))
#dev.off()

#CDismr3=ismr4

#pdf(sprintf("rslt/DESeq/GSEA_signature_comparisons_%s.pdf", SampleFrac), width=10, height=12)

for (i in 1:length(CDismr3)){
  x1=CDismr3[[i]]
  # find padj<0.1
  idx2=lapply(seq(3, ncol(x1), by=2), function(x) which(as.numeric(as.character(x1[ ,x]))<0.01))
  unList=table(unlist(idx2))
  #pickThese=names(unList)#[which(unList>1)]
  pickThese2=names(unList)[which(unList>1)]
  
  if (length(pickThese2)>100){
    temp=x1[as.numeric(pickThese2), seq(2, ncol(x1), by=2)]
    temp2=sapply(temp, function(x) as.numeric(as.character(x)))
    temp2=rowSds(abs(temp2))
    names(temp2)=pickThese2
    temp2=sort(temp2, decreasing = T)
    pickThese2=names(temp2)[1:100]
  }
  
  x23=x1[as.numeric(pickThese2), seq(2, ncol(x1), by=2)]
  
  x234=sapply(x23, function(x) as.numeric(as.character(x)))
  rownames(x234)=rownames(x23)
  
  par(oma=c(2,0,0,2))
  heatmap.2(x234, col=RdBu[11:1], scale="none", trace="none", main=sprintf("%s pathways %s fraction", 
                                                                           names(CDismr3)[i], SampleFrac))
}

#dev.off()


```

### DN samples

```{r}
SampleFrac="DN"

allDNGenes=unlist(DNCompSig)
allDNTab=table(allDNGenes)
DNGenesList=names(allDNTab)[which(allDNTab>1)]
#DNGenesList=DNGenesList[-grep("LOC", DNGenesList)]
#DNGenesList=DNGenesList[-grep("RGD", DNGenesList)]
# also, remove all samples which have a base mean less than 10
x1=DNComp4[[1]]$baseMean[match(DNGenesList, DNComp4[[1]]$Gene)]
x2=which(x1>25)
vstDN=vst(DNdds, blind=T)

DNddsTreatG$Comp8_v2=factor(paste(DNddsTreatG$Treatment, DNddsTreatG$Growth))

ColsideCols=DiffCols[DNddsTreatG$Comp8_v2]
ColsideColsB=hue_pal()(4)[DNddsTreatG$Treatment]
ColsideColsC=hue_pal()(2)[DNddsTreatG$Growth]

colTable=rbind(ColsideCols, ColsideColsB, ColsideColsC)
rownames(colTable)=c("Treat+growth", "Treatment", "growth")



#pdf(sprintf("rslt/DESeq/DEseq_comparisons_%s.pdf", SampleFrac), width=7, height=10)

sampleDist=dist(t(assay(vstDN)))
sampleDistMatrix <- as.matrix(sampleDist)
heatmap.plus(sampleDistMatrix, trace="none", scale="none", col=brewer.pal(11, "Blues")[11:1], ColSideColors = t(colTable), main=sprintf("Similarity between %s samples", SampleFrac))

# Take the top 150 variable genes 
# compute the sds for each genes
sds = rowSds(assay(vstDN))
means=rowMeans(assay(vstDN))
CVs=sds/means
CVs=CVs[-which(means<7)]
sds2=sort(CVs, decreasing = T)

heatmap.plus(assay(vstDN)[names(sds2)[1:200], ], trace="none", scale="row", col=RdBu[11:1], ColSideColors = t(colTable), main=sprintf("top 200 variable genes all %s samples", SampleFrac))


heatmap.plus(assay(vstDN)[DNGenesList[x2], ], trace="none", scale="row", col=RdBu[11:1], ColSideColors = t(colTable),
          main=sprintf("DEG at least 2 %s samples", SampleFrac))

# Make a table for these genes
FCvalues=matrix(NA, nrow=length(x2), ncol=length(DNComp4))

for (i in 1:length(DNComp4)){
  FCvalues[ ,i]=DNComp4[[i]]$log2FoldChange[match(DNGenesList[x2], DNComp4[[i]]$Gene)]
}

colnames(FCvalues)=names(DNComp4)
rownames(FCvalues)=DNGenesList[x2]


# for each list pick the top 50 diff genes FC>2 and baseman > 50
glist=NA
for (i in 1:length(DNComp4)){
  glist=c(glist, DNComp4[[i]]$Gene[which(DNComp4[[i]]$padj<0.1 & abs(DNComp4[[i]]$log2FoldChange)>2 & DNComp4[[i]]$baseMean>25)])
}

ColsideCols=SetCols[DNddsTreatG$Comp8]

glistmod=na.omit(unique(glist))
#glistmod=glistmod[-grep("LOC", glistmod)]
#glistmod=na.omit(glistmod[-grep("RGD", glistmod)])

heatmap.plus(assay(vstDN)[glistmod, ], trace="none", scale="row", col=RdBu[11:1], ColSideColors = t(colTable),
          main=sprintf("All DEG %s samples", SampleFrac))

# distance of the samples to one another
par(oma=c(2,0,0,0))
heatmap.2(FCvalues, trace="none", scale="row", col=RdBu[11:1], 
          main=sprintf("All comparisons %s, log2FC for each gene", SampleFrac))
#dev.off()

#DNismr3=ismr4

#pdf(sprintf("rslt/DESeq/GSEA_signature_comparisons_%s.pdf", SampleFrac), width=10, height=12)

for (i in 1:length(DNismr3)){
  x1=DNismr3[[i]]
  # find padj<0.1
  idx2=lapply(seq(3, ncol(x1), by=2), function(x) which(as.numeric(as.character(x1[ ,x]))<0.01))
  unList=table(unlist(idx2))
  #pickThese=names(unList)#[which(unList>1)]
  pickThese2=names(unList)[which(unList>1)]
  
  if (length(pickThese2)>100){
    temp=x1[as.numeric(pickThese2), seq(2, ncol(x1), by=2)]
    temp2=sapply(temp, function(x) as.numeric(as.character(x)))
    temp2=rowSds(abs(temp2))
    names(temp2)=pickThese2
    temp2=sort(temp2, decreasing = T)
    pickThese2=names(temp2)[1:100]
  }
  
  x23=x1[as.numeric(pickThese2), seq(2, ncol(x1), by=2)]
  
  x234=sapply(x23, function(x) as.numeric(as.character(x)))
  rownames(x234)=rownames(x23)
  
  par(oma=c(2,0,0,2))
  heatmap.2(x234, col=RdBu[11:1], scale="none", trace="none", main=sprintf("%s pathways %s fraction", 
                                                                           names(DNismr3)[i], SampleFrac))
}

#dev.off()

```

## Growing vs stable emphasis

Below, we focus specifically on the growing vs stable comparison in greater depth.
Here, we look at the 3 different fractions in greater depth and look at volcano plots of DEG and heatmaps of differential genes. Samples are colored according to whether they are growing or stable

### DN fraction

```{r}

#pdf(sprintf("rslt/DESeq/volcano_plots_differences_stable_vs_growing_%s.pdf", Sys.Date()), height=7, width=8)

DNa1=DNComp4$stable_vs_growing
DNa=DNComp4$stable_vs_growing[ which(DNComp4$stable_vs_growing$padj<0.1 & DNComp4$stable_vs_growing$baseMean>50 & abs(DNComp4$stable_vs_growing$log2FoldChange)>1.5), ]
with(DNa1, plot(log2FoldChange, -log10(padj), pch=20, main="Volcano plot: DN Stable (+) vs Growing (-)", cex=1.0, xlab=bquote(~Log[2]~fold~change), ylab=bquote(~-log[10]~Q~value)))
with(subset(DNa1, padj<0.1 & abs(log2FoldChange)>1.5), points(log2FoldChange, -log10(padj), pch=20, col="red", cex=0.5))
with(subset(DNa1, padj<0.1 & abs(log2FoldChange)>1.5), text(log2FoldChange+0.05, -log10(padj)+0.05, Gene, pch=20, col="red", cex=0.75))


heatmap.2(assay(vstDN)[ DNa$Gene, ], scale="row", trace="none", ColSideColors = as.character(as.numeric(vstDN$Growth)), col=RdBu[11:1], main="DN genes")
```

### CD fraction

```{r}
CDa1=CDComp4$stable_vs_growing
CDa=CDComp4$stable_vs_growing[ which(CDComp4$stable_vs_growing$padj<0.05 & CDComp4$stable_vs_growing$baseMean>100 & abs(CDComp4$stable_vs_growing$log2FoldChange)>2), ]
with(CDa1, plot(log2FoldChange, -log10(padj), pch=20, main="Volcano plot: CD Stable (+) vs Growing (-)", cex=1.0, xlab=bquote(~Log[2]~fold~change), ylab=bquote(~-log[10]~Q~value)))
with(subset(CDa1, padj<0.1 & abs(log2FoldChange)>1.5), points(log2FoldChange, -log10(padj), pch=20, col="red", cex=0.5))
with(subset(CDa1, padj<0.1 & abs(log2FoldChange)>1.5), text(log2FoldChange+0.05, -log10(padj)+0.05, Gene, pch=20, col="red", cex=0.75))

heatmap.2(assay(vstCD)[ CDa$Gene, ], scale="row", trace="none", ColSideColors = as.character(as.numeric(vstCD$Growth)), col=RdBu[11:1], main="CD genes")
```

### Ep Fraction

```{r}
Epa1=EpComp4$stable_vs_growing
Epa=EpComp4$stable_vs_growing[ which(EpComp4$stable_vs_growing$padj<0.1 & EpComp4$stable_vs_growing$baseMean>50 & abs(EpComp4$stable_vs_growing$log2FoldChange)>1.5), ]
with(Epa1, plot(log2FoldChange, -log10(padj), pch=20, main="Volcano plot: Ep Stable vs Growing", cex=1.0, xlab=bquote(~Log[2]~fold~change), ylab=bquote(~-log[10]~Q~value)))
with(subset(Epa1, padj<0.1 & abs(log2FoldChange)>1.5), points(log2FoldChange, -log10(padj), pch=20, col="red", cex=0.5))
with(subset(Epa1, padj<0.1 & abs(log2FoldChange)>1.5), text(log2FoldChange+0.05, -log10(padj)+0.1, Gene, pch=20, col="red", cex=0.75))

heatmap.2(assay(vstEp)[ Epa$Gene, ], scale="row", trace="none", ColSideColors = as.character(as.numeric(vstEp$Growth)), col=RdBu[11:1], main="Ep genes")


#dev.off()

```

### Specific pathways

The above plots suggest that there could be a different in growing and stable based on :

* inflammation
* Kras signalling
* MHC presentation
* checkpoint proteins

We can pull out the genes in these sets and visualise the relative expression in a heatmap in the DN, CD45 and Ep samples:

Note that red is growing and green is stable

```{r}
# library(org.Hs.eg.db)
# load("../anntotations/ListofGeneSets2.RData")

SetNamesc2=names(PathInc2) #names(ListGSC$c2List)
x1=grep("MHC", SetNamesc2)
GeneNames=unique(unlist(geneIds(PathInc2[x1[c(1, 4,5)]])))

#test1=mapIds(org.Hs.eg.db, GeneNames, 'SYMBOL','ENTREZID')
RatGeneNamesMHC=na.omit(unique(SymHum2Rat$RGD.symbol[match(GeneNames, SymHum2Rat$HGNC.symbol)]))

#RatGeneNamesKras=mapIds(org.Hs.eg.db, ListGSC$Hallmark$HALLMARK_KRAS_SIGNALING_UP, 'SYMBOL','ENTREZID')
GeneNames=unlist(geneIds(PathInH["HALLMARK_KRAS_SIGNALING_UP"]))
RatGeneNamesKras=na.omit(unique(SymHum2Rat$RGD.symbol[match(GeneNames, SymHum2Rat$HGNC.symbol)]))

GeneNames=unlist(geneIds(PathInH["HALLMARK_INFLAMMATORY_RESPONSE"]))
RatGeneNamesInf=na.omit(unique(SymHum2Rat$RGD.symbol[match(GeneNames, SymHum2Rat$HGNC.symbol)]))

CheckpointProt=unlist(ImmSuppAPC)
RatGeneNamesCheckpoint=na.omit(unique(SymHum2Rat$RGD.symbol[match(CheckpointProt, SymHum2Rat$HGNC.symbol)]))

# RatCP=read.delim("../anntotations/immune_Suppression.csv", sep="\t", stringsAsFactors = F)
# RatCP=na.omit(unique(SymHum2Rat$RGD.symbol[match(unique(unlist(RatCP)), SymHum2Rat$HGNC.symbol)]))
# 
# 
# RatMHC=read.csv("../anntotations/MHCloss.csv", header = F, stringsAsFactors = F)[ ,1]
# RatMHCb=na.omit(unique(SymHum2Rat$RGD.symbol[match(RatMHC, SymHum2Rat$HGNC.symbol)]))
# read in checkpoint proteins
```

#### DN samples:

Note the distribution of growing/stable in DN is: `r table(DNdds$Growth)`

```{r}
ColsideColsC=hue_pal()(2)[DNdds$Growth]

heatmap.2(assay(vstDN)[na.omit(match(RatGeneNamesMHC, rownames(vstDN))), ], col=RdBu[11:1], trace="none", scale="row", ColSideColors = ColsideColsC, main="MHC expression")

heatmap.2(assay(vstDN)[na.omit(match(RatGeneNamesKras, rownames(vstDN))), ], col=RdBu[11:1], trace="none", scale="row", ColSideColors = ColsideColsC, main="Kras expression")

heatmap.2(assay(vstDN)[na.omit(match(RatGeneNamesInf, rownames(vstDN))), ], col=RdBu[11:1], trace="none", scale="row", ColSideColors = ColsideColsC, main="Inflammatory response")

heatmap.2(assay(vstDN)[na.omit(match(RatGeneNamesCheckpoint, rownames(vstDN))), ], col=RdBu[11:1], trace="none", scale="row", ColSideColors = ColsideColsC, main="checkpoint proteins")

heatmap.2(assay(vstDN)[na.omit(match(MHCPres2Rat, rownames(vstDN))), ], col=RdBu[11:1], trace="none", scale="row", ColSideColors = ColsideColsC, main="MHC proteins")
```

#### CD45 samples:

Note the distribution of growing/stable in CD is: `r table(CDdds$Growth)`


```{r}
ColsideColsC=hue_pal()(2)[CDdds$Growth]

heatmap.2(assay(vstCD)[na.omit(match(RatGeneNamesMHC, rownames(vstCD))), ], col=RdBu[11:1], trace="none", scale="row", ColSideColors = ColsideColsC, main="MHC expression")

heatmap.2(assay(vstCD)[na.omit(match(RatGeneNamesKras, rownames(vstCD))), ], col=RdBu[11:1], trace="none", scale="row", ColSideColors = ColsideColsC, main="Kras expression")

heatmap.2(assay(vstCD)[na.omit(match(RatGeneNamesInf, rownames(vstCD))), ], col=RdBu[11:1], trace="none", scale="row", ColSideColors = ColsideColsC, main="Inflammatory response")

heatmap.2(assay(vstCD)[na.omit(match(RatGeneNamesCheckpoint, rownames(vstCD))), ], col=RdBu[11:1], trace="none", scale="row", ColSideColors = ColsideColsC, main="checkpoint proteins")

heatmap.2(assay(vstCD)[na.omit(match(MHCPres2Rat, rownames(vstCD))), ], col=RdBu[11:1], trace="none", scale="row", ColSideColors = ColsideColsC, main="MHC proteins")
```

#### Epithelial samples:

```{r}
ColsideColsC=hue_pal()(2)[Epdds$Growth]

heatmap.2(assay(vstEp)[na.omit(match(RatGeneNamesMHC, rownames(vstEp))), ], col=RdBu[11:1], trace="none", scale="row", ColSideColors = ColsideColsC, main="MHC expression")

heatmap.2(assay(vstEp)[na.omit(match(RatGeneNamesKras, rownames(vstEp))), ], col=RdBu[11:1], trace="none", scale="row", ColSideColors = ColsideColsC, main="Kras expression")

heatmap.2(assay(vstEp)[na.omit(match(RatGeneNamesInf, rownames(vstEp))), ], col=RdBu[11:1], trace="none", scale="row", ColSideColors = ColsideColsC, main="Inflammatory response")

heatmap.2(assay(vstEp)[na.omit(match(RatGeneNamesCheckpoint, rownames(vstEp))), ], col=RdBu[11:1], trace="none", scale="row", ColSideColors = ColsideColsC, main="checkpoint proteins")

heatmap.2(assay(vstEp)[na.omit(match(MHCPres2Rat, rownames(vstEp))), ], col=RdBu[11:1], trace="none", scale="row", ColSideColors = ColsideColsC, main="MHC proteins")
```



```{r, eval=F}
### comapre DN samples: PAM50 Lum vs Basal/Normal?

DNdds$PAM50_v2="ERplus"
DNdds$PAM50_v2[which(DNdds$PAM50%in%c("Basal", "Normal"))]="ERneg"
DNdds$PAM50_v2=factor(DNdds$PAM50_v2)
design(DNdds)=~PAM50_v2+factor(Batch)
DNddsP50=DESeq(DNdds)

DNresP50=results(DNddsP50)

x1=rownames(DNresP50)[which(abs(DNresP50$log2FoldChange)>2 & DNresP50$padj<0.05)]
```